---
title: "Water Level Data Summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
```

```{r Load Libraries, echo = FALSE, warning = FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(lubridate)
library(ggpubr)
library(tidytext)
```

### Introduction
This is a notebook to begin to investigate the water level sensor data for the FFAR project

### Data
Load all of the zentra information and extract the water level data from Michigan into a dataframe wl_dat 

```{r Load Zentra Datasets, echo=FALSE}
file <- "C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Water Cycling Indicators\\Data\\Soil Moisture Sensor Data\\Zentra_API\\Tall\\all_dat_4.csv" # set the working directory
all_dat <- read.csv(file)

all_dat_wl <- all_dat[all_dat$OutputName == "Water Level", ]
```

```{r Load depth to water key, echo = FALSE}
key_dat <- read.csv("C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Water Cycling Indicators\\Data\\Water_Level_Logger_Key.csv")

key_dat$Sensor_Depth_cm <- as.numeric(key_dat$Sensor_Depth_cm)
key_dat$Elevation_m <- as.numeric(key_dat$Elevation_m)
key_dat[is.na(key_dat$Sensor_Depth_cm),'Sensor_Depth_cm'] <- mean(key_dat$Sensor_Depth_cm,na.rm = TRUE)

```
Next, parse the "Site" column into Treatment, Block, and Triangle columns
```{r function to Load all ZENTRA from API, echo = FALSE, warning = FALSE}
wl_dat <- all_dat_wl
wl_dat$datetimeUTC <- as_datetime(wl_dat$datetime)
wl_dat$Name <- wl_dat$Site
wl_dat$Name <- gsub("^AD", "MI-AD", wl_dat$Name) # append MI site ID to these sites
wl_dat$Name <- gsub("^PR", "MI-PR", wl_dat$Name) # append MI site ID to these sites

wl_dat$Site <- sapply(strsplit(wl_dat$Name, "[_|-]"), "[[", 1)
wl_dat$Treatment <- sapply(strsplit(wl_dat$Name, "[_|-]"), "[[", 2)
wl_dat$Block <- sapply(strsplit(wl_dat$Name, "[_|-]"), "[[", 3)
wl_dat$Triangle <- sapply(strsplit(wl_dat$Name, "[_|-]"), "[[", 4)
wl_dat$Bl_Tri <- paste0(wl_dat$Block, "_", wl_dat$Triangle)

wl_dat$value <- as.numeric(wl_dat$value)

wl_dat$value[is.na(wl_dat$value)] <- 0
wl_dat$value[wl_dat$value < 25] <- 0 # assume negative or very small numbers (<2.5 cm) are not accurate and are best represented by 0

wl_dat <- merge(wl_dat, key_dat[, c("Name", "Sensor_Depth_cm", 'Elevation_m')], by = "Name")
wl_dat$dtw_m <- wl_dat$Sensor_Depth_cm/100 - wl_dat$value/1000
wl_dat$H <- 1.5 - wl_dat$dtw_m

wl_dat$H[wl_dat$value == 0] <- 0 # overwrite head as 0 which basically means unknown because water was at or below the sensor
wl_dat$H_elev <- wl_dat$Elevation_m - wl_dat$dtw_m 
```

```{r view dataframe}
head(wl_dat)
```

Then remove known bad stretches of data due to sensor failure 

```{r Clean up bad data stretches, warning = FALSE}
# wc_dat <- subset(wc_dat, !(Name == "WY_AD-11-12" & Port == 2))
# wc_dat <- wc_dat[!(wc_dat$Name == "MI-AD_9_1" & wc_dat$datetimeUTC < as_datetime("2022-8-15 19:00:00")), ]
# wc_dat <- wc_dat[!(wc_dat$Name == "MI-PR_9_9" & wc_dat$Port == 3 &
#                    wc_dat$datetimeUTC < as_datetime("2022-8-5 00:00:00") &
#                    wc_dat$datetimeUTC > as_datetime("2022-7-28 00:00:00")), ]
# wc_dat <- wc_dat[!(wc_dat$Name == "MI-PR_8_11" & wc_dat$Port == 3 &
#                    wc_dat$datetimeUTC < as_datetime("2022-7-16 18:00:00") &
#                    wc_dat$datetimeUTC > as_datetime("2022-7-16 12:00:00")), ]
# wc_dat <- wc_dat[!(wc_dat$Name == "RR_PR_4_7" & wc_dat$Port == 3 &
#                    wc_dat$datetimeUTC < as_datetime("2022-10-05 11:00:00") &
#                    wc_dat$datetimeUTC > as_datetime("2022-8-24 02:00:00")), ]
# wc_dat <- wc_dat[!(wc_dat$Name == "RR_PR_4_7" & wc_dat$Port == 3 &
#                    wc_dat$datetimeUTC < as_datetime("2022-12-07 13:00:00") &
#                    wc_dat$datetimeUTC > as_datetime("2022-11-04 12:00:00")), ]
```

```{r Make some  plotting functions, echo = FALSE}
get_wl_plot <- function(dat, t1, t2, tr){
      p1 <- ggplot() + geom_line(data = dat,
          aes(x = datetimeUTC, y = H, group = as.factor(Block), color = as.factor(Block)),
          size = 1, show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(paste0(dat$Site[1], " - ", tr), subtitle = paste0("Block ", dat$Block[1])) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
          xlab("Date") + ylab("Water Level (m)") +
          xlim(c(t1, t2)) + facet_grid(Block ~ Triangle)
      return(p1)
      }

get_wl_elev_plot <- function(dat, t1, t2, tr){
      p1 <- ggplot() + geom_line(data = dat,
          aes(x = datetimeUTC, y = H_elev, group = as.factor(Block), color = as.factor(Block)),
          size = 1, show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(paste0(dat$Site[1], " - ", tr), subtitle = paste0("Block ", dat$Block[1])) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
          xlab("Date") + ylab("Water Elevation (m)") +
          xlim(c(t1, t2)) + facet_grid(Block ~ Triangle)
      return(p1)
      }


get_wl_d_plot <- function(dat, t1, t2){
     p3 <- ggplot() + geom_line(data = dat, aes(x = datetimeUTC, y = H, color = Name), size = 1) +
     geom_line(data = dat, aes(x = datetimeUTC, y = sp_avg_H), color = 'black', linewidth = 2, show.legend = FALSE) +
     theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
     theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
     xlab("Date") + ylab("Water Level (m)") +
     xlim(c(t1, t2)) + facet_wrap(~ Treatment)
     return(p3)
}

get_wl_d_elev_plot <- function(dat, t1, t2){
     p3 <- ggplot() + geom_line(data = dat, aes(x = datetimeUTC, y = H_elev, color = Name), linewidth = 1, show.legend =TRUE) +
     # geom_line(data = dat, aes(x = datetimeUTC, y = sp_avg_H_elev), color = 'black', linewidth = 2, show.legend = FALSE) +
     theme_bw(base_size = 16) +
     theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "right") +
     xlab("Date") + ylab("Water Table Elevation (m)") +
     xlim(c(t1, t2)) 
     return(p3)
}

get_wl_mrd_plot <- function(dat, dat_mean, t1, t2){
    p4 <-ggplot() + geom_line(data = dat, aes(x = datetimeUTC, y = rd, color = Name), size = 1) +
     geom_hline(data = dat_mean, aes(yintercept = avg_rd, color = Name), linewidth = 1, alpha = 0.4) +
     geom_hline(yintercept = 0, color = "Black", linewidth = 2) +
     theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
     theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
     xlab("Date") + ylab("Relative Difference in Water Level (-)") +
     xlim(c(t1, t2)) +
     facet_wrap(~ Treatment)
     return(p4)
}

get_wl_mrd_elev_plot <- function(dat, dat_mean, t1, t2){
    p4 <-ggplot() + geom_line(data = dat, aes(x = datetimeUTC, y = rd_elev, color = Name), size = 1) +
     geom_hline(data = dat_mean, aes(yintercept = avg_rd, color = Name), linewidth = 1, alpha = 0.4) +
     geom_hline(yintercept = 0, color = "Black", linewidth = 2) +
     theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
     theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
     xlab("Date") + ylab("Relative Difference in Water Level (-)") +
     xlim(c(t1, t2)) +
     facet_wrap(~ Treatment)
     return(p4)
}

get_mrd_bar_plot <- function(dat){
     p2 <- ggplot(data = dat,
          aes(x = Bl_Tri, y = avg_rd, fill = as.factor(Name))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Mean Relative Difference in Water Level (-)") +
          facet_wrap(~ Treatment, nrow = 3, scale = "free_x")
          scale_fill_discrete(name = "Triangle")
     return(p2)
}

get_mrd_bar_plot2 <- function(dat){
     p2 <- ggplot(data = dat,
          aes(reorder_within(Bl_Tri, avg_rd, list(Block,Treatment)), avg_rd, fill = as.factor(Name))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Mean Relative Difference in Water Level") +
          scale_x_reordered() +
          facet_wrap(~ Treatment, nrow = 3, scale = "free_x")
     return(p2)
}

get_sd_bar_plot <- function(dat){
     p2 <- ggplot(data = dat,
          aes(x = Bl_Tri, y = sd_rd, fill = as.factor(Name))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("SD Relative Difference in Water Level") +
          facet_wrap(~ Treatment, nrow = 3, scale = "free_x")
          scale_fill_discrete(name = "Triangle")
     return(p2)
}

get_sd_bar_plot2 <- function(dat){
     p2 <- ggplot(data = dat,
          aes(reorder_within(Bl_Tri, sd_rd, list(Block,Treatment)), sd_rd, fill = as.factor(Name))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Standard Deviation of Relative Difference in Water Level") +
          scale_x_reordered() +
          facet_wrap(~ Treatment, nrow = 3, scale = "free_x")
     return(p2)
}
```

## Visualize all Water Content Data for each sensor individually 

```{r Plot each Block via a loop, warning = FALSE, fig.width=15,fig.height=5}
t1 <- as_datetime("2022-07-01 00:00:00")
t2 <- as_datetime("2023-5-30 00:00:00")

for (site in unique(wl_dat$Site)){
    for (block in unique(wl_dat$Block[wl_dat$Site == site])){
            tr <- "AD"
            dat <- subset(wl_dat, Site == site & Block == block & Treatment == tr)
            p1 <- get_wl_plot(dat, t1, t2, tr)
            tr <- "PR"
            dat <- subset(wl_dat, Site == site & Block == block & Treatment == tr)
            p2 <- get_wl_plot(dat, t1, t2, tr)
            p <- ggarrange(p1, p2, ncol = 2)
            print(p)
    }
}
```

## Investigate spatiotemporal stability and persistence 

Calculate the mean water content and the mean relative difference 
(relative to other sensors at that site and treatment)
of each sensor

For hourly data:

```{r Calculate the temporal stability with HOURLY data, warning = FALSE}
# Calculate the mean and standard deviation of each sensor relative to other sensors in that site/treatment pair
wl_dat_mean <- wl_dat %>% dplyr::group_by(Name, Site, Treatment, Block, Triangle) %>% 
               dplyr::summarize(avg_H = mean(H, na.rm = TRUE), sd_H = sd(H, na.rm = TRUE))

wl_dat_mrd <- wl_dat %>% dplyr::group_by(Site, Treatment, datetimeUTC) %>% 
               dplyr::mutate(sp_avg_H = mean(H, na.rm = TRUE), sp_sd_H = sd(H, na.rm = TRUE))

wl_dat_mrd$rd <- (wl_dat_mrd$H - wl_dat_mrd$sp_avg_H)/wl_dat_mrd$sp_avg_H

wl_dat_mrd_mean <- wl_dat_mrd %>% dplyr::group_by(Name, Site, Treatment, Block, Bl_Tri, Triangle) %>% 
               dplyr::summarize(avg_rd = mean(rd, na.rm = TRUE), sd_rd = sd(rd, na.rm = TRUE))

```

... and for daily data:

```{r Calculate the temporal stability with DAILY data, warning = FALSE}
# Reduce to daily data by extracting measurements from the noon hour only and repeat calculation
wl_dat_daily <- subset(wl_dat, hour(datetimeUTC) == 12)
wl_dat_daily_mean <- wl_dat_daily %>% dplyr::group_by(Name, Site, Treatment, Block, Triangle) %>% 
               dplyr::summarize(avg_H = mean(H, na.rm = TRUE), avg_H_elev = mean(H_elev, na.rm = TRUE), sd_H = sd(H, na.rm = TRUE))

wl_dat_daily_mrd <- wl_dat_daily %>% dplyr::group_by(Site, datetimeUTC) %>% 
               dplyr::mutate(sp_avg_H = mean(H, na.rm = TRUE), sp_avg_H_elev = mean(H_elev, na.rm = TRUE), sp_sd_H = sd(H, na.rm = TRUE))

wl_dat_daily_mrd$rd <- (wl_dat_daily_mrd$H - wl_dat_daily_mrd$sp_avg_H)/wl_dat_daily_mrd$sp_avg_H
wl_dat_daily_mrd$rd_elev <- (wl_dat_daily_mrd$H_elev - wl_dat_daily_mrd$sp_avg_H_elev)/wl_dat_daily_mrd$sp_avg_H_elev

wl_dat_daily_mrd_mean <- wl_dat_daily_mrd %>% dplyr::group_by(Name, Site, Treatment, Block, Bl_Tri, Triangle) %>% 
               dplyr::summarize(avg_rd = mean(rd, na.rm = TRUE), sd_rd = sd(rd, na.rm = TRUE))

```

### Water Level through time 
Grouped for each site/treatment pair with spatially averaged water content in black

```{r plot difference wl, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
for (site in unique(wl_dat$Site)){
     dat <- subset(wl_dat_daily_mrd, Site == site)
     p <- get_wl_d_plot(dat, t1, t2)
     print(p)
    }
```

```{r plot difference wl elev, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
for (site in unique(wl_dat$Site)){
     dat <- subset(wl_dat_daily_mrd, Site == site)
     p <- get_wl_d_elev_plot(dat, t1, t2)
     print(p)
    }
```

```{r plot difference wl elev block 8, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
for (site in unique(wl_dat$Site)){
     for (block in unique(wl_dat$Block)){
          # for (tr in c('AD','PR')){
               dat <- subset(wl_dat_daily_mrd, Site == site & Block == block)
               # p <- get_wl_d_elev_plot(dat, t1, t2)
               # print(p)
               pdf(paste0("C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Documents\\FFAR\\Conferences\\20230611_CUAHSI\\Poster\\Graphics\\WLxtime_",block,".pdf"), width=4, height=3)
               get_wl_d_elev_plot(dat, t1, t2)
               # print(p1)
               dev.off()
          }
     }
```

### Relative Difference of Water Level 
From the spatial average for each site/treatment pair (0 Line in black)

```{r plot mean relative difference wl, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
for (site in unique(wl_dat$Site)){
     dat <- subset(wl_dat_daily_mrd, Site == site)
     dat_mean <-  subset(wl_dat_daily_mrd_mean, Site == site)
     p <- get_wl_mrd_plot(dat, dat_mean, t1, t2)
     print(p)
    }
```


## Temporal mean & standard deviation of relative difference in water level of each sensor 

```{r Plot temporal stability by site for DAILY data unRanked, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
# for (site in unique(wl_dat_daily_mrd_mean$Site)){
#      dat <- subset(wl_dat_daily_mrd_mean, Site == site)
#      p1 <- get_mrd_bar_plot(dat)
#      p2 <- get_sd_bar_plot(dat)
#      p <- ggarrange(p1, p2, ncol = 2)
#      print(p)
#     }
```

### Ranked
```{r Plot temporal stability by site for DAILY data RANKED, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
for (site in unique(wl_dat_daily_mrd_mean$Site)){
     dat <- subset(wl_dat_daily_mrd_mean, Site == site)
     p1 <- get_mrd_bar_plot2(dat)
     p2 <- get_sd_bar_plot2(dat)
     p <- ggarrange(p1, p2, ncol = 2)
     print(p)
    }
```