---
title: "Saturo Data Summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
```

```{r Load Libraries, echo = FALSE, warning = FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(lubridate)
library(ggpubr)
library(tidytext)
library(readxl)
```

### Introduction
This is a notebook to begin to investigate the raw saturo data for the FFAR project

### Data
Load all of the spreadsheet information 

```{r Load Raw Saturo Files, echo=FALSE, warning = FALSE}
path2dir <- "C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Water Cycling Indicators\\Data\\Infiltration Data\\"
folders <- dir(path2dir)
folders <- subset(folders, grepl("Raw", folders))

folder <- folders[1]
file <- dir(paste0(path2dir, folder))[1]
sat_dat <- list()

for (folder in folders){
     for (file in dir(paste0(path2dir, folder))){
          path2file <- paste0(path2dir, folder, "\\", file)
          season <- strsplit(folder, "[ |-]")[[1]][7]
          name <- paste0(file, "-", season)
          sat_dat[[name]]  <- as.data.frame(read_excel(path = path2file, sheet = "Raw Data"))
          sat_dat[[name]]$season <- season
          sat_dat[[name]]$FileName <- file
          if (!is.na(match("Flux (cm/s)", colnames(sat_dat[[name]])))){
               sat_dat[[name]][,"Flux (cm/hr)"] <- sat_dat[[name]][,"Flux (cm/s)"]*3600 # convert cm/s to cm/hr
          }
          if (!is.na(match("Volume (mL/s)", colnames(sat_dat[[name]])))){
               sat_dat[[name]][,"Volume (mL/hr)"] <- sat_dat[[name]][,"Volume (mL/s)"]*3600 # convert cm/s to cm/hr
          }

          summary_dat  <- as.data.frame(read_excel(path = path2file, sheet = "Summary"))
          sat_dat[[name]]$Kfs <- as.numeric(summary_dat[21, 2])
          sat_dat[[name]]$KfsErr <- as.numeric(summary_dat[22, 2])
          if (!is.na(match("Kfs (cm/s)", summary_dat[21, 1]))){
               sat_dat[[name]]$Kfs <- sat_dat[[name]]$Kfs*3600 # convert cm/s to cm/hr
               sat_dat[[name]]$KfsErr <- sat_dat[[name]]$KfsErr*3600 # convert cm/s to cm/hr
          }
     }
}

sat_dat_melt <- melt(sat_dat, id.vars = c("season", "FileName", "Time (min)", "Water Level (cm)", "Pressure (cm)", "Flux (cm/hr)", "Volume (mL/hr)", "Kfs", "KfsErr"))
```


Next, parse the "Site" column into Treatment, Block, and Triangle columns
```{r parse Saturo data columns, echo = FALSE, warning = FALSE}
sat_dat_melt$FileName <- gsub("^22-", "", sat_dat_melt$FileName) # append MI site ID to these sites
sat_dat_melt$FileName <- gsub("^OK-", "", sat_dat_melt$FileName) # append MI site ID to these sites

sat_dat_melt$Site <- sapply(strsplit(sat_dat_melt$FileName, "[_|-]"), "[[", 1)
sat_dat_melt$Treatment <- sapply(strsplit(sat_dat_melt$FileName, "[_|-]"), "[[", 2)
sat_dat_melt$Block <- sapply(strsplit(sat_dat_melt$FileName, "[_|-]"), "[[", 3)

# if site was WY you need to flip the treatment and block labels
bl_tmp <- sat_dat_melt[sat_dat_melt$Site == "WY", "Block"]
sat_dat_melt[sat_dat_melt$Site == "WY", "Block"] <- sat_dat_melt[sat_dat_melt$Site == "WY", "Treatment"]
sat_dat_melt[sat_dat_melt$Site == "WY", "Treatment"] <- bl_tmp

sat_dat_melt$Triangle <- sapply(strsplit(sat_dat_melt$FileName, "[ |.|(|)|_|-]"), "[[", 4)
sat_dat_melt$Bl_Tri <- paste0(sat_dat_melt$Block, "_", sat_dat_melt$Triangle)
```

```{r view dataframe}
head(sat_dat_melt)
```

Then remove known bad stretches of data due to sensor failure 

```{r Clean up bad data stretches, warning = FALSE}
# wc_dat <- subset(wc_dat, !(Name == "WY_AD-11-12" & Port == 2))
# wc_dat <- wc_dat[!(wc_dat$Name == "MI-AD_9_1" & wc_dat$datetimeUTC < as_datetime("2022-8-15 19:00:00")), ]
# wc_dat <- wc_dat[!(wc_dat$Name == "MI-PR_9_9" & wc_dat$Port == 3 &
#                    wc_dat$datetimeUTC < as_datetime("2022-8-5 00:00:00") &
#                    wc_dat$datetimeUTC > as_datetime("2022-7-28 00:00:00")), ]
# wc_dat <- wc_dat[!(wc_dat$Name == "MI-PR_8_11" & wc_dat$Port == 3 &
#                    wc_dat$datetimeUTC < as_datetime("2022-7-16 18:00:00") &
#                    wc_dat$datetimeUTC > as_datetime("2022-7-16 12:00:00")), ]
# wc_dat <- wc_dat[!(wc_dat$Name == "RR_PR_4_7" & wc_dat$Port == 3 &
#                    wc_dat$datetimeUTC < as_datetime("2022-10-05 11:00:00") &
#                    wc_dat$datetimeUTC > as_datetime("2022-8-24 02:00:00")), ]
# wc_dat <- wc_dat[!(wc_dat$Name == "RR_PR_4_7" & wc_dat$Port == 3 &
#                    wc_dat$datetimeUTC < as_datetime("2022-12-07 13:00:00") &
#                    wc_dat$datetimeUTC > as_datetime("2022-11-04 12:00:00")), ]
```

```{r Make some  plotting functions, echo = FALSE}
get_flux_plot <- function(dat, tr){
      p1 <- ggplot() + geom_point(data = dat,
          aes(x = `Time (min)`, y = `Flux (cm/hr)`, group = as.factor(season), color = as.factor(season)),
          size = 1, show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(paste0(dat$Site[1], " - ", tr), subtitle = paste0("Block ", dat$Block[1])) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
          xlab("Time (min)") + ylab("Flux (cm/hr)") +
          facet_wrap(FileName ~ Triangle, ncol = 3, scale = "free_y")
      return(p1)
      }


get_pressure_plot <- function(dat, tr){
      p1 <- ggplot() + geom_point(data = dat,
          aes(x = `Time (min)`, y = `Pressure (cm)`, group = as.factor(season), color = as.factor(season)),
          size = 1, show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(paste0(dat$Site[1], " - ", tr), subtitle = paste0("Block ", dat$Block[1])) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
          xlab("Time (min)") + ylab("Pressure (cm)") +
          facet_wrap(FileName ~ Triangle, ncol = 3, scale = "free_y")
      return(p1)
      }

get_KFS_plot <- function(dat, tr){
      p1 <- ggplot() + geom_point(data = dat,
          aes(x = `Time (min)`, y = `Flux (cm/hr)`/`Pressure (cm)`, group = as.factor(season), color = as.factor(season)),
          size = 1, show.legend = FALSE) +
          geom_hline(data = dat, aes(yintercept = Kfs, color = as.factor(season)), linewidth = 1, alpha = 0.4, show.legend = FALSE) + 
          theme_bw(base_size = 16) + ggtitle(paste0(dat$Site[1], " - ", tr), subtitle = paste0("Block ", dat$Block[1])) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
          xlab("Time (min)") + ylab("Flux/Pressure") +
          facet_wrap(FileName ~ Triangle, ncol = 3, scale = "free_y")
      return(p1)
      }
# get_wc_d_plot <- function(dat, t1, t2){
#      p3 <- ggplot() + geom_point(data = dat, aes(x = datetimeUTC, y = value, color = Name), size = 1) +
#      geom_line(data = dat, aes(x = datetimeUTC, y = sp_avg_value), color = 'black', linewidth = 2, show.legend = FALSE) +
#      theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
#      theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
#      xlab("Date") + ylab("Water Content (cm^3/cm^3)") +
#      xlim(c(t1, t2)) + facet_grid(FileName ~ Block)
#      return(p3)
# }


# get_wc_mrd_plot <- function(dat, dat_mean, t1, t2){
#     p4 <-ggplot() + geom_point(data = dat, aes(x = datetimeUTC, y = rd, color = Name), size = 1) +
#      geom_hline(data = dat_mean, aes(yintercept = avg_rd, color = Name), linewidth = 1, alpha = 0.4) +
#      geom_hline(yintercept = 0, color = "Black", linewidth = 2) +
#      theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
#      theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
#      xlab("Date") + ylab("Relative Difference in Water Content (cm^3/cm^3)") +
#      xlim(c(t1, t2)) +
#      facet_grid(Port ~ Treatment)
#      return(p4)
# }

# get_mrd_bar_plot <- function(dat){
#      p2 <- ggplot(data = dat,
#           aes(x = Bl_Tri, y = avg_rd, fill = as.factor(Name))) +
#           geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
#           theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
#           theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
#           xlab("Site") + ylab("Mean Relative Difference in Water Content") +
#           facet_wrap(Port ~ Treatment, nrow = 3, scale = "free_x")
#           scale_fill_discrete(name = "Triangle")
#      return(p2)
# }

# get_mrd_bar_plot2 <- function(dat){
#      p2 <- ggplot(data = dat,
#           aes(reorder_within(Bl_Tri, avg_rd, list(Port,Treatment)), avg_rd, fill = as.factor(Name))) +
#           geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
#           theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
#           theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
#           xlab("Site") + ylab("Mean Relative Difference in Water Content") +
#           scale_x_reordered() +
#           facet_wrap(Port ~ Treatment, nrow = 3, scale = "free_x")
#      return(p2)
# }

# get_sd_bar_plot <- function(dat){
#      p2 <- ggplot(data = dat,
#           aes(x = Bl_Tri, y = sd_rd, fill = as.factor(Name))) +
#           geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
#           theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
#           theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
#           xlab("Site") + ylab("SD Relative Difference in Water Content") +
#           facet_wrap(Port ~ Treatment, nrow = 3, scale = "free_x")
#           scale_fill_discrete(name = "Triangle")
#      return(p2)
# }

# get_sd_bar_plot2 <- function(dat){
#      p2 <- ggplot(data = dat,
#           aes(reorder_within(Bl_Tri, sd_rd, list(Port,Treatment)), sd_rd, fill = as.factor(Name))) +
#           geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
#           theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
#           theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
#           xlab("Site") + ylab("Standard Deviation of Relative Difference in Water Content") +
#           scale_x_reordered() +
#           facet_wrap(Port ~ Treatment, nrow = 3, scale = "free_x")
#      return(p2)
# }
```

## Visualize all the Saturo Data for each experiment

### Plot raw flux data (cm/hr)

```{r Plot Flux cm/s data, echo = FALSE, warning = FALSE, fig.width=15,fig.height=10}
for (site in unique(sat_dat_melt$Site)){
    for (block in unique(sat_dat_melt$Block[sat_dat_melt$Site == site])){
            tr <- "AD"
            dat <- subset(sat_dat_melt, Site == site & Block == block & Treatment == tr)
            p1 <- get_flux_plot(dat, tr)
            tr <- "PR"
            dat <- subset(sat_dat_melt, Site == site & Block == block & Treatment == tr)
            p2 <- get_flux_plot(dat, tr)
            p <- ggarrange(p1, p2, ncol = 2)
            print(p)
    }
}
```


### Plot raw pressure data (cm)

```{r Plot Pressure mL/s data, echo = FALSE, warning = FALSE, fig.width=15,fig.height=10}
for (site in unique(sat_dat_melt$Site)){
    for (block in unique(sat_dat_melt$Block[sat_dat_melt$Site == site])){
            tr <- "AD"
            dat <- subset(sat_dat_melt, Site == site & Block == block & Treatment == tr)
            p1 <- get_pressure_plot(dat, tr)
            tr <- "PR"
            dat <- subset(sat_dat_melt, Site == site & Block == block & Treatment == tr)
            p2 <- get_pressure_plot(dat, tr)
            p <- ggarrange(p1, p2, ncol = 2)
            print(p)
    }
}
```

### Plot raw flux/pressure 

```{r Plot Flux/Pressure data, echo = FALSE, warning = FALSE, fig.width=15,fig.height=10}
for (site in unique(sat_dat_melt$Site)){
    for (block in unique(sat_dat_melt$Block[sat_dat_melt$Site == site])){
            tr <- "AD"
            dat <- subset(sat_dat_melt, Site == site & Block == block & Treatment == tr)
            p1 <- get_KFS_plot(dat, tr)
            tr <- "PR"
            dat <- subset(sat_dat_melt, Site == site & Block == block & Treatment == tr)
            p2 <- get_KFS_plot(dat, tr)
            p <- ggarrange(p1, p2, ncol = 2)
            print(p)
    }
}
```

## Investigate spatiotemporal stability and persistence 

Calculate the mean water content and the mean relative difference 
(relative to other sensors at that site and treatment)
of each sensor

For hourly data:

```{r Calculate the temporal stability with HOURLY data, warning = FALSE}
# Calculate the mean and standard deviation of each sensor relative to other sensors in that site/treatment pair
# wc_dat_mean <- wc_dat %>% dplyr::group_by(Name, Site, Treatment, Block, Triangle, Port) %>% 
#                dplyr::summarize(avg_value = mean(value, na.rm = TRUE), sd_value = sd(value, na.rm = TRUE))

# wc_dat_mrd <- wc_dat %>% dplyr::group_by(Site, Treatment, Port, datetimeUTC) %>% 
#                dplyr::mutate(sp_avg_value = mean(value, na.rm = TRUE), sp_sd_value = sd(value, na.rm = TRUE))

# wc_dat_mrd$rd <- (wc_dat_mrd$value - wc_dat_mrd$sp_avg_value)/wc_dat_mrd$sp_avg_value

# wc_dat_mrd_mean <- wc_dat_mrd %>% dplyr::group_by(Name, Site, Treatment, Block, Bl_Tri, Triangle, Port) %>% 
#                dplyr::summarize(avg_rd = mean(rd, na.rm = TRUE), sd_rd = sd(rd, na.rm = TRUE))

```

... and for daily data:

```{r Calculate the temporal stability with DAILY data, warning = FALSE}
# Reduce to daily data by extracting measurements from the noon hour only and repeat calculation
# wc_dat_daily <- subset(wc_dat, hour(datetimeUTC) == 12)
# wc_dat_daily_mean <- wc_dat_daily %>% dplyr::group_by(Name, Site, Treatment, Block, Triangle, Port) %>% 
#                dplyr::summarize(avg_value = mean(value, na.rm = TRUE), sd_value = sd(value, na.rm = TRUE))

# wc_dat_daily_mrd <- wc_dat_daily %>% dplyr::group_by(Site, Treatment, Port, datetimeUTC) %>% 
#                dplyr::mutate(sp_avg_value = mean(value, na.rm = TRUE), sp_sd_value = sd(value, na.rm = TRUE))

# wc_dat_daily_mrd$rd <- (wc_dat_daily_mrd$value - wc_dat_daily_mrd$sp_avg_value)/wc_dat_daily_mrd$sp_avg_value

# wc_dat_daily_mrd_mean <- wc_dat_daily_mrd %>% dplyr::group_by(Name, Site, Treatment, Block, Bl_Tri, Triangle, Port) %>% 
#                dplyr::summarize(avg_rd = mean(rd, na.rm = TRUE), sd_rd = sd(rd, na.rm = TRUE))

```

### Water Content through time 
Grouped for each site/treatment pair with spatially averaged water content in black

```{r plot difference wc, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
# t1 <- as_datetime("2022-07-01 00:00:00")
# t2 <- as_datetime("2023-4-6 00:00:00")

# for (site in unique(wc_dat$Site)){
#      dat <- subset(wc_dat_daily_mrd, Site == site)
#      p <- get_wc_d_plot(dat, t1, t2)
#      print(p)
#     }
```

### Relative Difference of Water Content 
From the spatial average for each site/treatment pair (0 Line in black)

```{r plot mean relative difference wc, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
# t1 <- as_datetime("2022-07-01 00:00:00")
# t2 <- as_datetime("2023-4-6 00:00:00")

# for (site in unique(wc_dat$Site)){
#      dat <- subset(wc_dat_daily_mrd, Site == site)
#      dat_mean <-  subset(wc_dat_daily_mrd_mean, Site == site)
#      p <- get_wc_mrd_plot(dat, dat_mean, t1, t2)
#      print(p)
#     }
```


## Temporal mean & standard deviation of relative difference in water content of each sensor 

### Unranked, ordered by Block 
```{r Plot temporal stability by site for DAILY data unRanked, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
# for (site in unique(wc_dat_daily_mrd_mean$Site)){
#      dat <- subset(wc_dat_daily_mrd_mean, Site == site)
#      p1 <- get_mrd_bar_plot(dat)
#      p2 <- get_sd_bar_plot(dat)
#      p <- ggarrange(p1, p2, ncol = 2)
#      print(p)
#     }
```

### Ranked
```{r Plot temporal stability by site for DAILY data RANKED, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
# for (site in unique(wc_dat_daily_mrd_mean$Site)){
#      dat <- subset(wc_dat_daily_mrd_mean, Site == site)
#      p1 <- get_mrd_bar_plot2(dat)
#      p2 <- get_sd_bar_plot2(dat)
#      p <- ggarrange(p1, p2, ncol = 2)
#      print(p)
#     }
```