---
title: "Saturo Data Summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
```

```{r Load Libraries, echo = FALSE, warning = FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(lubridate)
library(ggpubr)
library(tidytext)
library(readxl)
```

### Introduction
This is a notebook to begin to investigate the raw saturo data for the FFAR project

### Data
Load all of the spreadsheet information 

```{r Load Raw Saturo Files, echo=FALSE, warning = FALSE}
path2dir <- "C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Water Cycling Indicators\\Data\\Infiltration Data\\"
folders <- dir(path2dir)
folders <- subset(folders, grepl("Raw", folders) & !grepl("Producer",folders))

folder <- folders[1]
file <- dir(paste0(path2dir, folder))[1]
sat_dat <- list()

for (folder in folders){
     for (file in dir(paste0(path2dir, folder))){
          path2file <- paste0(path2dir, folder, "\\", file)
          season <- strsplit(folder, "[ |-]")[[1]][7]
          name <- paste0(file, "-", season)
          sat_dat[[name]]  <- as.data.frame(read_excel(path = path2file, sheet = "Raw Data"))
          sat_dat[[name]]$season <- season
          sat_dat[[name]]$year <- strsplit(folder, "[ |-]")[[1]][6]
          sat_dat[[name]]$FileName <- file
          if (!is.na(match("Flux (cm/s)", colnames(sat_dat[[name]])))){
               sat_dat[[name]][,"Flux (cm/hr)"] <- sat_dat[[name]][,"Flux (cm/s)"]*3600 # convert cm/s to cm/hr
          }
          if (!is.na(match("Volume (mL/s)", colnames(sat_dat[[name]])))){
               sat_dat[[name]][,"Volume (mL/hr)"] <- sat_dat[[name]][,"Volume (mL/s)"]*3600 # convert cm/s to cm/hr
          }

          summary_dat  <- as.data.frame(read_excel(path = path2file, sheet = "Summary"))
          sat_dat[[name]]$Serial <- summary_dat[11,2]
          sat_dat[[name]]$Kfs <- as.numeric(summary_dat[21, 2])
          sat_dat[[name]]$KfsErr <- as.numeric(summary_dat[22, 2])
          if (!is.na(match("Kfs (cm/s)", summary_dat[21, 1]))){
               sat_dat[[name]]$Kfs <- sat_dat[[name]]$Kfs*3600 # convert cm/s to cm/hr
               sat_dat[[name]]$KfsErr <- sat_dat[[name]]$KfsErr*3600 # convert cm/s to cm/hr
          }
     }
}

sat_dat_melt <- melt(sat_dat, id.vars = c("season", "year", "FileName", "Serial", "Time (min)", "Water Level (cm)", "Pressure (cm)", "Flux (cm/hr)", "Volume (mL/hr)", "Kfs", "KfsErr"))
```


Next, parse the "Site" column into Treatment, Block, and Triangle columns
```{r parse Saturo data columns, echo = FALSE, warning = FALSE}
sat_dat_melt$FileName <- gsub("^22-", "", sat_dat_melt$FileName) # Drop 22 from filename
sat_dat_melt$FileName <- gsub("^23-", "", sat_dat_melt$FileName) # Drop 23 from filename
sat_dat_melt$FileName <- gsub("^_ 23-", "", sat_dat_melt$FileName) # Drop _ 23 from filename
sat_dat_melt$FileName <- gsub("SP-", "", sat_dat_melt$FileName) 
sat_dat_melt$FileName <- gsub("^OK-", "", sat_dat_melt$FileName) 

sat_dat_melt$Site <- sapply(strsplit(sat_dat_melt$FileName, "[_|-]"), "[[", 1)
sat_dat_melt$Treatment <- sapply(strsplit(sat_dat_melt$FileName, "[_|-]"), "[[", 2)
sat_dat_melt$Block <- sapply(strsplit(sat_dat_melt$FileName, "[_|-]"), "[[", 3)

# if site was WY in 2022 you need to flip the treatment and block labels
bl_tmp <- sat_dat_melt[sat_dat_melt$Site == "WY" & sat_dat_melt$year == '2022', "Block"]
sat_dat_melt[sat_dat_melt$Site == "WY" & sat_dat_melt$year == '2022', "Block"] <- sat_dat_melt[sat_dat_melt$Site == "WY" & sat_dat_melt$year == '2022', "Treatment"]
sat_dat_melt[sat_dat_melt$Site == "WY" & sat_dat_melt$year == '2022', "Treatment"] <- bl_tmp

sat_dat_melt$Triangle <- sapply(strsplit(sat_dat_melt$FileName, "[ |.|(|)|_|-]"), "[[", 4)
sat_dat_melt$Bl_Tri <- paste0(sat_dat_melt$Block, "_", sat_dat_melt$Triangle)
```

```{r view dataframe}
head(sat_dat_melt)
```

```{r Make some  plotting functions, echo = FALSE}
get_flux_plot <- function(dat, tr){
      p1 <- ggplot() + geom_line(data = dat,
          aes(x = `Time (min)`, y = `Flux (cm/hr)`, group = FileName, color = FileName),
          size = 1, show.legend = TRUE) +
          theme_bw(base_size = 16) + ggtitle(paste0(dat$Site[1], " - ", tr), subtitle = paste0("Block ", dat$Block[1])) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
          xlab("Time (min)") + ylab("Flux (cm/hr)") +
          facet_wrap(Triangle ~ season, ncol = 2, scale = "free_y")
      return(p1)
      }


get_pressure_plot <- function(dat){
      p1 <- ggplot() + geom_line(data = dat,
          aes(x = `Time (min)`, y = `Pressure (cm)`, group = FileName, color = FileName),
          size = 1, show.legend = FALSE) + ggtitle(paste0(dat$Site[1], " - ", dat$season[1], ' ', dat$year[1])) + 
          theme_bw(base_size = 16) + 
          theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
          xlab("Time (min)") + ylab("Pressure (cm)") +
          facet_wrap(~Serial, nrow = 3, scale = "free_y")
      return(p1)
      }

get_KFS_plot <- function(dat, tr){
      p1 <- ggplot() + geom_point(data = dat,
          aes(x = `Time (min)`, y = `Flux (cm/hr)`/`Pressure (cm)`, group = FileName, color = FileName),
          size = 1, show.legend = TRUE) + 
          geom_hline(data = dat, aes(yintercept = Kfs, color = FileName), linewidth = 1, alpha = 0.4, show.legend = FALSE) + 
          theme_bw(base_size = 16) + ggtitle(paste0(dat$Site[1], " - ", tr), subtitle = paste0("Block ", dat$Block[1])) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
          xlab("Time (min)") + ylab("Flux/Pressure") +
          facet_wrap(Triangle ~ season, ncol = 2, scale = "free_y")
      return(p1)
      }

get_KFS_boxplot_byTri <- function(dat){
    p3 <- ggplot(data = dat, aes(x = reorder_within(Triangle,Kfs,Block), y = Kfs, fill = interaction(season,year))) +
        geom_boxplot() +
        xlab("Site") + ylab("Saturated K (cm/hr)") +
        theme_bw(base_size = 16) + ggtitle(paste0(dat$Site[1], " - ", tr), subtitle = paste0("Block ", dat$Block[1])) +
        theme(axis.text.x=element_text(angle=45,hjust=1)) + scale_fill_discrete(name = "Season and year") +
        facet_wrap(~Block, ncol = 4, scale = "free_x")
    return(p3)
    } 

get_flux_boxplot_byTri <- function(dat){
    p3 <- ggplot(data = dat, aes(x = reorder_within(Triangle,avg_flux,Block), y = avg_flux, fill = interaction(season,year))) +
        geom_boxplot() +
        xlab("Site") + ylab("Avg Flux (cm/hr)") +
        theme_bw(base_size = 16) + ggtitle(paste0(dat$Site[1], " - ", tr), subtitle = paste0("Block ", dat$Block[1])) +
        theme(axis.text.x=element_text(angle=45,hjust=1)) + scale_fill_discrete(name = "Season and year") +
        facet_wrap(~Block, ncol = 4, scale = "free_x")
    return(p3)
    } 

get_fluxvsksat_plot <- function(dat){
      p1 <- ggplot(data = dat,
          aes(x = Kfs, y = avg_flux, group = interaction(season,year), color = interaction(season,year), shape = Serial),
          size = 3, stroke = 2, show.legend = TRUE) + geom_point() + scale_shape_manual(values=1:9) + 
          geom_smooth(method = "lm") + stat_regline_equation(aes(label = ..rr.label..)) +
          theme_bw(base_size = 16) + # ggtitle(paste0(dat$Site[1], " - ", tr), subtitle = paste0("Block ", dat$Block[1])) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
          xlab("Ksat (cm/hr)") + ylab("avg Flux cm/hr") #+
      return(p1)
}

get_fluxvsksat_plot_all <- function(dat){
      p1 <- ggplot(data = dat,
          aes(x = log10(Kfs), y = log10(avg_flux)), show.legend = TRUE) + 
          geom_point(aes(color = Site, shape = season), size = 3) + 
          geom_smooth(method = "lm") + stat_regline_equation(aes(label = ..rr.label..)) +
          theme_bw(base_size = 16) + # ggtitle(paste0(dat$Site[1], " - ", tr), subtitle = paste0("Block ", dat$Block[1])) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
          xlab("Log Saturated K (cm/hr)") + ylab("Log avg Flux (cm/hr)") #+
        #   facet_wrap(~ Site, ncol = 2, scale = "free")
      return(p1)
      }
```

## Visualize all the Saturo Data for each experiment


### Plot raw pressure data by serial number (cm)

```{r Plot Pressure data, echo = FALSE, warning = FALSE, fig.width=15,fig.height=10}
for (site in unique(sat_dat_melt$Site)){
    for (yr in unique(sat_dat_melt$year)){
        dat <- subset(sat_dat_melt, Site == site & year == yr)
        for (seas in unique(dat$season)){
            p1 <- get_pressure_plot(subset(dat, season == seas))
            print(p1)
        }
    }
}
```

Then remove known bad stretches of data due to pressure issues 

```{r Clean up bad data stretches, warning = FALSE}
sat_dat_clean <- sat_dat_melt
sat_dat_clean <- subset(sat_dat_clean, !(Serial == 'SAT00002868' & (year == "2022" | year == "2023")))

sat_dat_clean <- subset(sat_dat_clean, !(FileName == "WY-PR-14-12 (2).xlsx" | FileName == "WY-PR-14-12 (3).xlsx"))
sat_dat_clean <- subset(sat_dat_clean, !(FileName == "MI-PR-8-11 (2).xlsx" | FileName == "MI-PR-8-11.xlsx"))

sat_dat_clean <- subset(sat_dat_clean, !(FileName == "MI-PR-8-2 (1).xlsx"))
sat_dat_clean <- subset(sat_dat_clean, !(FileName == "RR-AD-6-6 (2).xlsx"))

```

```{r check locations with issues , echo = FALSE, warning = FALSE, fig.width=15,fig.height=10}
site = "WY"
yr = "2023"
seas = "Spring"
# ser = "SAT00002801"
dat <- subset(sat_dat_clean, Site == site & year == yr & season == seas & Treatment == "PR" & Block == 11 & Triangle == 12)
p1 <- ggplot() + geom_line(data = dat,
          aes(x = `Time (min)`, y = `Flux (cm/hr)`, group = FileName, color = FileName),
          size = 1, show.legend = FALSE) + 
          theme_bw(base_size = 16) + 
          theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
          xlab("Time (min)") + ylab("Pressure (cm)") +
          facet_wrap(~FileName)
p1
```

```{r Plot Pressure after cleaning , echo = FALSE, warning = FALSE, fig.width=15,fig.height=10}
for (site in unique(sat_dat_clean$Site)){
    for (yr in unique(sat_dat_clean$year)){
        dat <- subset(sat_dat_clean, Site == site & year == yr)
        for (seas in unique(dat$season)){
            p1 <- get_pressure_plot(subset(dat, season == seas))
            print(p1)
        }
    }
}
```

### Plot raw flux data (cm/hr)

```{r Plot Flux cm/s data, echo = FALSE, warning = FALSE, fig.width=15,fig.height=10}
# for (site in unique(sat_dat_melt$Site)){
#     for (block in unique(sat_dat_melt$Block[sat_dat_melt$Site == site])){
#             tr <- "AD"
#             dat <- subset(sat_dat_melt, Site == site & Block == block & Treatment == tr)
#             p1 <- get_flux_plot(dat, tr)
#             tr <- "PR"
#             dat <- subset(sat_dat_melt, Site == site seas& Block == block & Treatment == tr)
#             p2 <- get_flux_plot(dat, tr)
#             p <- ggarrange(p1, p2, ncol = 2)
#             print(p)
#     }
# }
```

### Plot box plots for each site and treatment
```{r Plot boxplots of KFS data, echo = FALSE, warning = FALSE, fig.width=15,fig.height=5}
# for (site in unique(sat_dat_clean$Site)){
#     for (block in unique(sat_dat_clean$Block[sat_dat_clean$Site == site])){
#             tr <- "AD"
#             dat <- subset(sat_dat_clean, Site == site & Block == block & Treatment == tr)
#             p1 <- get_KFS_boxplot_byTri(dat)
#             tr <- "PR"
#             dat <- subset(sat_dat_clean, Site == site & Block == block & Treatment == tr)
#             p2 <- get_KFS_boxplot_byTri(dat)
#             p <- ggarrange(p1, p2, ncol = 2)
#             print(p)
#     }
# }
```


### Plot flux vs ksat at all 4 intensive hubs and all seasons
```{r Plot Flux cm/s vs ksat data, echo = FALSE, warning = FALSE, fig.width=15,fig.height=10}
dat <- sat_dat_clean %>% dplyr::group_by(L1, Kfs, Serial, season, year, Site, Treatment, Block, Triangle) %>% 
               dplyr::summarize(avg_flux = mean(`Flux (cm/hr)`, na.rm = TRUE))
# dat <- dat[!(dat$Site == 'WY' & dat$Kfs > 0.3*60), ]
p <- get_fluxvsksat_plot_all(dat[dat$Kfs > 0,]) # only use +K values to look at this graph
print(p)

```

```{r Plot boxplots of avg flux data, echo = FALSE, warning = FALSE, fig.width=15,fig.height=5}
dat <- sat_dat_clean %>% dplyr::group_by(L1, Serial, Kfs, season, year, Site, Block, Treatment, Triangle) %>% 
               dplyr::summarize(avg_flux = mean(`Flux (cm/hr)`, na.rm = TRUE))

for (site in unique(dat$Site)){
    for (block in unique(dat$Block[dat$Site == site])){
            tr <- "AD"
            p1 <- get_flux_boxplot_byTri(dat[dat$Site == site & dat$Block == block & dat$Treatment == tr,])
            tr <- "PR"
            p2 <- get_flux_boxplot_byTri(dat[dat$Site == site & dat$Block == block & dat$Treatment == tr,])
            p <- ggarrange(p1, p2, ncol = 2)
            print(p)
    }
}
```

```{r repeat with ksat calculated from slope of flux, echo = FALSE, warning = FALSE, fig.width=15,fig.height=5}
dat <- sat_dat_clean[sat_dat_clean$variable == 'Flux (cm/s)' & sat_dat_clean$`Time (min)` > 45,] %>% dplyr::group_by(L1, Kfs, season, year, Site, Block, Treatment, Triangle)  %>% 
               dplyr::mutate(cumsum_flux = cumsum(`Flux (cm/hr)`), slope_flux = coefficients(lm(cumsum_flux ~ `Time (min)`))[2])


dat$Kfs <- dat$slope_flux*0.5 - 0.012

for (site in unique(dat$Site)){
    for (block in unique(dat$Block[dat$Site == site])){
            tr <- "AD"
            p1 <- get_KFS_boxplot_byTri(dat[dat$Site == site & dat$Block == block & dat$Treatment == tr,])
            tr <- "PR"
            p2 <- get_KFS_boxplot_byTri(dat[dat$Site == site & dat$Block == block & dat$Treatment == tr,])
            p <- ggarrange(p1, p2, ncol = 2)
            print(p)
    }
}
```
