---
title: "Water Content Data Summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
```

```{r Load Libraries, echo = FALSE, warning = FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(lubridate)
library(ggpubr)
library(tidytext)
```

### Introduction
This is a notebook to begin to investigate the soil water content sensor data for the FFAR project

### Data
Load all of the zentra information and extract the water content data into a dataframe wc_dat 

```{r Load Zentra Datasets from producer sites, echo=FALSE}
file <- "C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Water Cycling Indicators\\Data\\Soil Moisture Sensor Data\\Zentra_API\\Tall\\prod_dat_2023.csv" # set the working directory
prod_dat <- read.csv(file)

prod_dat_wc <- prod_dat[prod_dat$OutputName == "Water Content", ]
```

```{r Load depth to water key, echo = FALSE}
# key_dat <- read.csv("C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Water Cycling Indicators\\Data\\Water_Level_Logger_Key.csv")

# key_dat$Sensor_Depth_cm <- as.numeric(key_dat$Sensor_Depth_cm)
# key_dat$Elevation_m <- as.numeric(key_dat$Elevation_m)
# key_dat[is.na(key_dat$Sensor_Depth_cm),'Sensor_Depth_cm'] <- mean(key_dat$Sensor_Depth_cm,na.rm = TRUE)

```

Next, parse the "Site" column into Block, and Triangle columns
```{r function to Load all ZENTRA from API, echo = FALSE, warning = FALSE}
wc_dat <- prod_dat_wc
wc_dat$datetimeUTC <- as_datetime(wc_dat$datetime)
wc_dat$Name <- wc_dat$Site

wc_dat$Site <- sapply(strsplit(wc_dat$Name, "[_|-]"), "[[", 1)
wc_dat$Block <- sapply(strsplit(wc_dat$Name, "[_|-]"), "[[", 2)
wc_dat$Triangle <- sapply(strsplit(wc_dat$Name, "[_|-]"), "[[", 3)
wc_dat$Bl_Tri <- paste0(wc_dat$Block, "_", wc_dat$Triangle)

wc_dat$value <- as.numeric(wc_dat$value)

# exclude all data readings above 0.55 and na values
wc_dat <- subset(wc_dat, !is.na(value) & value < 0.55)
```

```{r convert water content to storage, echo=FALSE}
# this is the calculation for wyoming only
wc_dat$S <- wc_dat$value*10
wc_dat$S[wc_dat$Port == 3] <- wc_dat$S[wc_dat$Port == 3]*4.096

# at WY-13-15 we only were able to install 2 sensors, at 10 and 20 cm
wc_dat$S[wc_dat$Name == 'WY-13-15' & wc_dat$Port == 2] <- wc_dat$S[wc_dat$Name == 'WY-13-15' & wc_dat$Port == 2]*5.096

wc_dat_totS <- wc_dat %>% dplyr::group_by(Site, Name, datetimeUTC) %>% 
               dplyr::mutate(totS = sum(S))
wc_dat_totS$totS_in <- wc_dat_totS$totS*0.39371

```

```{r Set plotting window}
t1 <- as_datetime("2023-09-01 00:00:00")
t2 <- as_datetime("2023-09-30 00:00:00")
```

```{r plot water storage through time, echo=FALSE, warning = FALSE, fig.width=7,fig.height=5}
p1 <- ggplot() + geom_point(data = wc_dat_totS,
          aes(x = datetimeUTC, y = totS_in, group = as.factor(Bl_Tri), color = as.factor(Bl_Tri)),
          size = 1, show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle('WY Producer Sites') +
          theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
          xlab("Date") + ylab("Soil water storage upper 2 feet of soil (in)") +
          xlim(c(t1, t2)) + ylim(c(0,10)) + facet_wrap(~Block, ncol = 4)
p1
```

```{r view dataframe}
# head(wc_dat)

# id_df <- wc_dat[,c("Name","Treatment","Block","Triangle")]
# id_df_out <- unique(id_df)
# id_df_out <- id_df_out[order(id_df_out$Name),]
# write.table(id_df_out, "C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Water Cycling Indicators\\Database Design\\Triangle_List.csv", sep = ",", row.names = FALSE, quote = FALSE) 

```

Then remove known bad stretches of data due to sensor failure 

```{r Clean up bad data stretches, warning = FALSE}

```



```{r Make some  plotting functions, echo = FALSE}
get_wc_plot <- function(dat, t1, t2, Block){
      p1 <- ggplot() + geom_point(data = dat,
          aes(x = datetimeUTC, y = value, group = as.factor(Port), color = as.factor(Port)),
          size = 1, show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(paste0(dat$Site[1], " - ", Block), subtitle = paste0("Block ", dat$Block[1])) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
          xlab("Date") + ylab("Water Content (cm^3/cm^3)") +
          xlim(c(t1, t2)) + facet_grid(Port ~ Triangle)
      return(p1)
      }


get_wc_d_plot <- function(dat, t1, t2){
     p3 <- ggplot() + geom_point(data = dat, aes(x = datetimeUTC, y = value, color = Name), size = 1) +
     geom_line(data = dat, aes(x = datetimeUTC, y = sp_avg_value), color = 'black', linewidth = 2, show.legend = FALSE) +
     theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
     theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
     xlab("Date") + ylab("Water Content (cm^3/cm^3)") +
     xlim(c(t1, t2)) + facet_wrap(~Port, nrow = 3)
     return(p3)
}

# get_wc_d_plot_poster <- function(dat, t1, t2){
#      p3 <- ggplot() + geom_point(data = dat, aes(x = datetimeUTC, y = value, color = Block), size = 1) +
#      geom_point(data = dat, aes(x = datetimeUTC, y = sp_avg_value), color = 'black', size = 1, show.legend = FALSE) +
#      theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
#      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#      xlab("Date") + ylab("Water Content (cm^3/cm^3)") +
#      xlim(c(t1, t2)) + facet_wrap(~ Treatment, nrow =2)
#      return(p3)
# }


get_wc_mrd_plot <- function(dat, dat_mean, t1, t2){
    p4 <-ggplot() + geom_point(data = dat, aes(x = datetimeUTC, y = rd, color = Name), size = 1) +
     geom_hline(data = dat_mean, aes(yintercept = avg_rd, color = Name), linewidth = 1, alpha = 0.4) +
     geom_hline(yintercept = 0, color = "Black", linewidth = 2) +
     theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
     theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
     xlab("Date") + ylab("Relative Difference in Water Content (cm^3/cm^3)") +
     xlim(c(t1, t2)) +
     facet_wrap(~Port, nrow = 3)
     return(p4)
}

get_mrd_bar_plot <- function(dat){
     p2 <- ggplot(data = dat,
          aes(x = Bl_Tri, y = avg_rd, fill = as.factor(Block))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Mean Relative Difference in Water Content") +
          facet_wrap(Port ~ Treatment, nrow = 3, scale = "free_x")
          scale_fill_discrete(name = "Triangle")
     return(p2)
}

get_mrd_bar_plot2 <- function(dat){
     p2 <- ggplot(data = dat,
          aes(reorder_within(Bl_Tri, avg_rd, list(Port,Treatment)), avg_rd, fill = as.factor(Block))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Mean Relative Difference in Water Content") +
          scale_x_reordered() +
          facet_wrap(Port ~ Treatment, nrow = 3, scale = "free_x")
     return(p2)
}

get_mrd_bar_plot3 <- function(dat){
     p2 <- ggplot(data = dat,
          aes(reorder_within(Bl_Tri, avg_rd, list(Treatment)), avg_rd, fill = as.factor(Block))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Mean Relative Difference in Water Content") +
          scale_x_reordered() +
          facet_wrap(~Treatment, nrow = 1, scale = "free_x")
     return(p2)
}

get_mrd_bar_plot4 <- function(dat){
     p2 <- ggplot(data = dat,
           aes(reorder_within(Bl_Tri, Elevation_m, list(Treatment)), avg_rd)) +
           geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
           geom_col_pattern(aes(pattern = Treatment, fill = Block), # <- make this an interaction
          #  colour = "black",
           pattern_fill = "black",
           pattern_angle = 0,
           pattern_density = 0.02,
           pattern_spacing = 0.02) +
           scale_pattern_manual(values=c('crosshatch', 'circle')) +
           theme_bw(base_size = 16) 
           + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Mean Relative Difference in Water Content") +
          scale_x_reordered() 
     return(p2)
}

get_mrd_bar_plot5 <- function(dat){
     p2 <- ggplot(data = dat,
           aes(reorder_within(Bl_Tri, SWS, list(Treatment)), avg_rd)) +
           geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
           geom_col_pattern(aes(pattern = Treatment, fill = Block), # <- make this an interaction
          #  colour = "black",
           pattern_fill = "black",
           pattern_angle = 0,
           pattern_density = 0.02,
           pattern_spacing = 0.02) +
           scale_pattern_manual(values=c('crosshatch', 'circle')) +
           theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Mean Relative Difference in Water Content") +
          scale_x_reordered() 
     return(p2)
}



get_sd_bar_plot <- function(dat){
     p2 <- ggplot(data = dat,
          aes(x = Bl_Tri, y = sd_rd, fill = as.factor(Name))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("SD Relative Difference in Water Content") +
          facet_wrap(Port ~ Treatment, nrow = 3, scale = "free_x")
          scale_fill_discrete(name = "Triangle")
     return(p2)
}

get_sd_bar_plot2 <- function(dat){
     p2 <- ggplot(data = dat,
          aes(reorder_within(Bl_Tri, sd_rd, list(Port,Treatment)), sd_rd, fill = as.factor(Name))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Standard Deviation of Relative Difference in Water Content") +
          scale_x_reordered() +
          facet_wrap(Port ~ Treatment, nrow = 3, scale = "free_x")
     return(p2)
}

get_sd_bar_plot3 <- function(dat){
     p2 <- ggplot(data = dat,
          aes(reorder_within(Bl_Tri, sd_rd, list(Treatment)), sd_rd, fill = as.factor(Name))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Standard Deviation of Relative Difference in Water Content") +
          scale_x_reordered() +
          facet_wrap(~Treatment, nrow = 1, scale = "free_x")
     return(p2)
}
```

## Visualize all Water Content Data for each sensor individually 

```{r Plot each Block via a loop, echo=FALSE, warning = FALSE, fig.width=15,fig.height=5}
# t1 <- as_datetime("2022-07-01 00:00:00")
# t2 <- as_datetime("2023-5-30 00:00:00")

# for (site in unique(wc_dat$Site)){
#     for (block in unique(wc_dat$Block[wc_dat$Site == site])){
#             tr <- "AD"
#             dat <- subset(wc_dat, Site == site & Block == block & Treatment == tr)
#             p1 <- get_wc_plot(dat, t1, t2, tr)
#             tr <- "PR"
#             dat <- subset(wc_dat, Site == site & Block == block & Treatment == tr)
#             p2 <- get_wc_plot(dat, t1, t2, tr)
#             p <- ggarrange(p1, p2, ncol = 2)
#             print(p)
#     }
# }
```

## Investigate spatiotemporal stability and persistence 

Calculate the mean water content and the mean relative difference 
(relative to other sensors at that site and treatment)
of each sensor

For hourly data:

```{r Calculate the temporal stability with HOURLY data, warning = FALSE}
# Calculate the mean and standard deviation of each sensor relative to other sensors in that site/treatment pair
wc_dat_mean <- wc_dat %>% dplyr::group_by(Name, Site, Block, Triangle, Port) %>% 
               dplyr::summarize(avg_value = mean(value, na.rm = TRUE), sd_value = sd(value, na.rm = TRUE))

wc_dat_mrd <- wc_dat %>% dplyr::group_by(Site, Port, datetimeUTC) %>% 
               dplyr::mutate(sp_avg_value = mean(value, na.rm = TRUE), sp_sd_value = sd(value, na.rm = TRUE))

wc_dat_mrd$rd <- (wc_dat_mrd$value - wc_dat_mrd$sp_avg_value)/wc_dat_mrd$sp_avg_value

wc_dat_mrd_mean <- wc_dat_mrd %>% dplyr::group_by(Name, Site, Block, Bl_Tri, Triangle, Port) %>% 
               dplyr::summarize(avg_rd = mean(rd, na.rm = TRUE), sd_rd = sd(rd, na.rm = TRUE))

```

... and for daily data:

```{r Calculate the temporal stability with DAILY data, warning = FALSE}
# Reduce to daily data by extracting measurements from the noon hour only and repeat calculation
wc_dat_daily <- subset(wc_dat, hour(datetimeUTC) == 12 & minute(datetimeUTC) == 0)

wc_dat_daily_mean <- wc_dat_daily %>% dplyr::group_by(Name, Site,  Block, Triangle, Port) %>% 
               dplyr::summarize(avg_value = mean(value, na.rm = TRUE), sd_value = sd(value, na.rm = TRUE))

wc_dat_daily_mrd <- wc_dat_daily %>% dplyr::group_by(Site, Port, datetimeUTC) %>% 
               dplyr::mutate(sp_avg_value = mean(value, na.rm = TRUE), sp_sd_value = sd(value, na.rm = TRUE))

wc_dat_daily_mrd$rd <- (wc_dat_daily_mrd$value - wc_dat_daily_mrd$sp_avg_value)/wc_dat_daily_mrd$sp_avg_value

wc_dat_daily_mrd_mean <- wc_dat_daily_mrd %>% dplyr::group_by(Name, Site, Block, Bl_Tri, Triangle, Port) %>% 
               dplyr::summarize(avg_rd = mean(rd, na.rm = TRUE), sd_rd = sd(rd, na.rm = TRUE))

wc_dat_daily_mrd_mean_alldepth <- wc_dat_daily_mrd %>% dplyr::group_by(Name, Site, Block, Bl_Tri, Triangle) %>% 
               dplyr::summarize(avg_rd = mean(rd, na.rm = TRUE), sd_rd = sd(rd, na.rm = TRUE))

```

### Water Content through time 
Grouped for each site/block pair with spatially averaged water content in black

```{r plot difference wc, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
for (site in unique(wc_dat_daily$Site)){
        dat <- subset(wc_dat_daily_mrd, Site == site)
        p <- get_wc_d_plot(dat, t1, t2)
        print(p)
    }
```

### Relative Difference of Water Content 
From the spatial average for each site/treatment pair (0 Line in black)

```{r plot mean relative difference wc, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
for (site in unique(wc_dat_daily$Site)){
     dat <- subset(wc_dat_daily_mrd, Site == site)
     dat_mean <-  subset(wc_dat_daily_mrd_mean, Site == site)
     p <- get_wc_mrd_plot(dat, dat_mean, t1, t2)
     print(p)
    }
```


## Temporal mean & standard deviation of relative difference in water content of each sensor 

```{r Plot temporal stability by site for DAILY data unRanked, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
# for (site in unique(wc_dat_daily_mrd_mean$Site)){
#      dat <- subset(wc_dat_daily_mrd_mean, Site == site)
#      p1 <- get_mrd_bar_plot(dat)
#      p2 <- get_sd_bar_plot(dat)
#      p <- ggarrange(p1, p2, ncol = 2)
#      print(p)
#     }
```

### Ranked
```{r Plot temporal stability by site & depth for DAILY data RANKED, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
for (site in unique(wc_dat_daily_mrd_mean$Site)){
     dat <- subset(wc_dat_daily_mrd_mean, Site == site)
     p1 <- get_mrd_bar_plot2(dat)
     # p2 <- get_sd_bar_plot2(dat)
     # p <- ggarrange(p1, p2, ncol = 2)
     print(p1)
    }
```

### Ranked and aggregated by depth
```{r Plot temporal stability by site DAILY data RANKED, warning = FALSE, echo = FALSE, fig.width=15,fig.height=4}
for (site in unique(wc_dat_daily_mrd_mean_alldepth$Site)){
     dat <- subset(wc_dat_daily_mrd_mean_alldepth, Site == site)
     p1 <- get_mrd_bar_plot3(dat)
     # p2 <- get_sd_bar_plot3(dat)
     # p <- ggarrange(p1, p2, ncol = 2)
     print(p1)
    }
```


### MI sites only Ranked by elevation and aggregated by depth
```{r Plot temporal stability by elevation DAILY data RANKED, warning = FALSE, echo = FALSE, fig.width=15,fig.height=4}
# add elevation data
pdf("C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Documents\\FFAR\\Conferences\\20230611_CUAHSI\\Poster\\Graphics\\diffWCxElev.pdf", width=6, height=4)
wc_dat3 <- merge(wc_dat_daily_mrd_mean_alldepth, key_dat[, c("Name", "Sensor_Depth_cm", 'Elevation_m','SWS')], by = "Name")
get_mrd_bar_plot4(wc_dat3)
# print(p1)
dev.off()
```

```{r Plot temporal stability by SWS DAILY data RANKED, warning = FALSE, echo = FALSE, fig.width=15,fig.height=4}
# add elevation data
pdf("C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Documents\\FFAR\\Conferences\\20230611_CUAHSI\\Poster\\Graphics\\diffWCxSWS.pdf", width=6, height=4)
get_mrd_bar_plot5(wc_dat3)
# print(p1)
dev.off()
```


```{r plot difference wc, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
pdf("C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Documents\\FFAR\\Conferences\\20230611_CUAHSI\\Poster\\Graphics\\WCxtime15cm.pdf", width=7, height=5)

t1 <- as_datetime("2022-08-01 00:00:00")
t2 <- as_datetime("2022-12-15 00:00:00")
site ='MI'
dat <- subset(wc_dat_daily_mrd, Site == site & Port == 1)
get_wc_d_plot_poster(dat, t1, t2)
# print(p)
dev.off()
```
