---
title: "Water Content Data Summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
```

```{r Load Libraries, echo = FALSE, warning = FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(lubridate)
library(ggpubr)
library(tidytext)
```

### Introduction
This is a notebook to begin to investigate the soil water content sensor data for the FFAR project

### Data
Load all of the zentra information and extract the water content data into a dataframe wc_dat 

```{r Load Zentra Datasets, echo=FALSE}
file <- "C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Water Cycling Indicators\\Data\\Soil Moisture Sensor Data\\Zentra_API\\Tall\\all_dat_2023.csv" # set the working directory
all_dat <- read.csv(file)

all_dat_wc <- all_dat[all_dat$OutputName == "Water Content", ]
```

Next, parse the "Site" column into Treatment, Block, and Triangle columns
```{r function to Load all ZENTRA from API, echo = FALSE, warning = FALSE}
wc_dat <- all_dat_wc
wc_dat$datetimeUTC <- as_datetime(wc_dat$datetime)
wc_dat$Name <- wc_dat$Site
wc_dat$Name <- gsub("^AD", "MI-AD", wc_dat$Name) # append MI site ID to these sites
wc_dat$Name <- gsub("^PR", "MI-PR", wc_dat$Name) # append MI site ID to these sites

wc_dat[wc_dat$Name == "WY_AD-10-3N", "Name"] <- "WY_AD-10-3" # drop the N off the label for WY_AD-10-3 so it will plot together with the other

wc_dat$Site <- sapply(strsplit(wc_dat$Name, "[_|-]"), "[[", 1)
wc_dat$Treatment <- sapply(strsplit(wc_dat$Name, "[_|-]"), "[[", 2)
wc_dat$Block <- sapply(strsplit(wc_dat$Name, "[_|-]"), "[[", 3)
wc_dat$Triangle <- sapply(strsplit(wc_dat$Name, "[_|-]"), "[[", 4)
wc_dat$Bl_Tri <- paste0(wc_dat$Block, "_", wc_dat$Triangle)

wc_dat$Port[wc_dat$Port == 5] <- 3 # rename the port 5 sensor as 3

wc_dat$value <- as.numeric(wc_dat$value)
```

```{r view dataframe, echo =FALSE}
# head(wc_dat)

# id_df <- wc_dat[,c("Name","Treatment","Block","Triangle")]
# id_df_out <- unique(id_df)
# id_df_out <- id_df_out[order(id_df_out$Name),]
# write.table(id_df_out, "C:\\Users\\kuhlalex\\OneDrive - Michigan State University\\Water Cycling Indicators\\Database Design\\Triangle_List.csv", sep = ",", row.names = FALSE, quote = FALSE) 

```

Then remove known bad stretches of data due to sensor failure 

```{r Clean up bad data stretches, warning = FALSE}
wc_dat <- subset(wc_dat, !(Name == "WY_AD-11-12" & Port == 2))
wc_dat <- wc_dat[!(wc_dat$Name == "MI-AD_9_1" & wc_dat$datetimeUTC < as_datetime("2022-8-15 19:00:00")), ]
wc_dat <- wc_dat[!(wc_dat$Name == "MI-PR_9_9" & wc_dat$Port == 3 &
                   wc_dat$datetimeUTC < as_datetime("2022-8-5 00:00:00") &
                   wc_dat$datetimeUTC > as_datetime("2022-7-28 00:00:00")), ]
wc_dat <- wc_dat[!(wc_dat$Name == "MI-PR_8_11" & wc_dat$Port == 3 &
                   wc_dat$datetimeUTC < as_datetime("2022-7-16 18:00:00") &
                   wc_dat$datetimeUTC > as_datetime("2022-7-16 12:00:00")), ]
wc_dat <- wc_dat[!(wc_dat$Name == "RR_PR_4_7" & wc_dat$Port == 3 &
                   wc_dat$datetimeUTC < as_datetime("2022-10-05 11:00:00") &
                   wc_dat$datetimeUTC > as_datetime("2022-8-24 02:00:00")), ]
wc_dat <- wc_dat[!(wc_dat$Name == "RR_PR_4_7" & wc_dat$Port == 3 &
                   wc_dat$datetimeUTC < as_datetime("2022-12-07 13:00:00") &
                   wc_dat$datetimeUTC > as_datetime("2022-11-04 12:00:00")), ]
```

```{r Make some  plotting functions, echo = FALSE}
get_wc_plot <- function(dat, t1, t2, tr){
      p1 <- ggplot() + geom_point(data = dat,
          aes(x = datetimeUTC, y = value, group = as.factor(Port), color = as.factor(Port)),
          size = 1, show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(paste0(dat$Site[1], " - ", tr), subtitle = paste0("Block ", dat$Block[1])) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
          xlab("Date") + ylab("Water Content (cm^3/cm^3)") +
          xlim(c(t1, t2)) + facet_grid(Port ~ Triangle)
      return(p1)
      }


get_wc_d_plot <- function(dat, t1, t2){
     p3 <- ggplot() + geom_line(data = dat, aes(x = datetimeUTC, y = value, color = Name), linewith = 1) +
     geom_line(data = dat, aes(x = datetimeUTC, y = sp_avg_value), color = 'black', linewidth = 2, show.legend = FALSE) +
     theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
     theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
     xlab("Date") + ylab("Water Content (cm^3/cm^3)") +
     xlim(c(t1, t2)) + facet_wrap(~Port, nrow = 3)
     return(p3)
}


get_wc_mrd_plot <- function(dat, dat_mean, t1, t2){
    p4 <-ggplot() + geom_line(data = dat, aes(x = datetimeUTC, y = rd*100, color = Name), linewidth = 1) +
     geom_hline(data = dat_mean, aes(yintercept = avg_rd, color = Name), linewidth = 1, alpha = 0.4) +
     geom_hline(yintercept = 0, color = "Black", linewidth = 2) +
     theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
     theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
     xlab("Date") + ylab("% Relative Difference in Water Content") +
     xlim(c(t1, t2)) +
    facet_wrap(~Port, nrow = 3)
     return(p4)
}

get_mrd_bar_plot <- function(dat){
     p2 <- ggplot(data = dat,
          aes(x = Bl_Tri, y = avg_rd*100, fill = as.factor(Treatment))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Mean % Relative Difference in Water Content") +
          facet_wrap(~Port, nrow = 3, scale = "free_x")
          scale_fill_discrete(name = "Triangle")
     return(p2)
}

get_mrd_bar_plot2 <- function(dat){
     p2 <- ggplot(data = dat,
          aes(reorder_within(Bl_Tri, avg_rd, list(Port,Treatment)), avg_rd*100, fill = as.factor(Treatment))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Mean % Relative Difference in Water Content") +
          scale_x_reordered() +
          facet_wrap(~Port, nrow = 3, scale = "free_x")
     return(p2)
}

get_sd_bar_plot <- function(dat){
     p2 <- ggplot(data = dat,
          aes(x = Bl_Tri, y = sd_rd, fill = as.factor(Treatment))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("SD Relative Difference in Water Content") +
          facet_wrap(~Port, nrow = 3, scale = "free_x")
          scale_fill_discrete(name = "Triangle")
     return(p2)
}

get_sd_bar_plot2 <- function(dat){
     p2 <- ggplot(data = dat,
          aes(reorder_within(Bl_Tri, sd_rd, list(Port,Treatment)), sd_rd, fill = as.factor(Treatment))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Standard Deviation of Relative Difference in Water Content") +
          scale_x_reordered() +
          facet_wrap(~Port, nrow = 3, scale = "free_x")
     return(p2)
}

get_corr_bar_plot2 <- function(dat){
     p2 <- ggplot(data = dat,
          aes(reorder_within(Bl_Tri, avg_corr, list(Port,Treatment)), avg_corr, fill = as.factor(Treatment))) +
          geom_bar(position = "dodge", stat = "identity", show.legend = FALSE) +
          theme_bw(base_size = 16) + ggtitle(dat$Site[1]) +
          theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          xlab("Site") + ylab("Mean Correlation in Water Content") +
          scale_x_reordered() +
          facet_wrap(~Port, nrow = 3, scale = "free_x")
     return(p2)
}
```

## Visualize all Water Content Data for each sensor individually 

```{r Plot each Block via a loop, echo = FALSE, warning = FALSE, fig.width=15,fig.height=5}
# t1 <- as_datetime("2022-07-01 00:00:00")
# t2 <- as_datetime("2023-4-6 00:00:00")

# for (site in unique(wc_dat$Site)){
#     for (block in unique(wc_dat$Block[wc_dat$Site == site])){
#             tr <- "AD"
#             dat <- subset(wc_dat, Site == site & Block == block & Treatment == tr)
#             p1 <- get_wc_plot(dat, t1, t2, tr)
#             tr <- "PR"
#             dat <- subset(wc_dat, Site == site & Block == block & Treatment == tr)
#             p2 <- get_wc_plot(dat, t1, t2, tr)
#             p <- ggarrange(p1, p2, ncol = 2)
#             print(p)
#     }
# }
```

## Investigate spatiotemporal stability and persistence 

Calculate the mean water content and the mean relative difference 
(relative to other sensors at that site and treatment)
of each sensor

For hourly data:

```{r Calculate the temporal stability with HOURLY data, warning = FALSE}
# Calculate the mean and standard deviation of each sensor relative to other sensors in that site/treatment pair
wc_dat_mean <- wc_dat %>% dplyr::group_by(Name, Site, Treatment, Block, Triangle, Port) %>% 
               dplyr::summarize(avg_value = mean(value, na.rm = TRUE), sd_value = sd(value, na.rm = TRUE))

wc_dat_mrd <- wc_dat %>% dplyr::group_by(Site, Port, datetimeUTC) %>% 
               dplyr::mutate(sp_avg_value = mean(value, na.rm = TRUE), sp_sd_value = sd(value, na.rm = TRUE))

wc_dat_mrd$rd <- (wc_dat_mrd$value - wc_dat_mrd$sp_avg_value)/wc_dat_mrd$sp_avg_value

wc_dat_mrd_mean <- wc_dat_mrd %>% dplyr::group_by(Name, Site, Treatment, Block, Bl_Tri, Triangle, Port) %>% 
               dplyr::summarize(avg_rd = mean(rd, na.rm = TRUE), sd_rd = sd(rd, na.rm = TRUE))

```

... and for daily data:

```{r Calculate the temporal stability with DAILY data, warning = FALSE}
# Reduce to daily data by extracting measurements from the noon hour only and repeat calculation
wc_dat_daily <- subset(wc_dat, hour(datetimeUTC) == 12)
wc_dat_daily_mean <- wc_dat_daily %>% dplyr::group_by(Name, Site, Treatment, Block, Triangle, Port) %>% 
               dplyr::summarize(avg_value = mean(value, na.rm = TRUE), sd_value = sd(value, na.rm = TRUE))

wc_dat_daily_mrd <- wc_dat_daily %>% dplyr::group_by(Site, Port, datetimeUTC) %>% 
               dplyr::mutate(sp_avg_value = mean(value, na.rm = TRUE), sp_sd_value = sd(value, na.rm = TRUE))

wc_dat_daily_mrd$rd <- (wc_dat_daily_mrd$value - wc_dat_daily_mrd$sp_avg_value)/wc_dat_daily_mrd$sp_avg_value

wc_dat_daily_mrd_mean <- wc_dat_daily_mrd %>% dplyr::group_by(Name, Site, Treatment, Block, Bl_Tri, Triangle, Port) %>% 
               dplyr::summarize(avg_rd = mean(rd, na.rm = TRUE), sd_rd = sd(rd, na.rm = TRUE))

```

### Water Content through time 
Grouped for each site/treatment pair with spatially averaged water content in black

```{r plot difference wc, warning = FALSE, echo = FALSE, fig.width=8,fig.height=5}
t1 <- as_datetime("2023-03-01 00:00:00")
t2 <- as_datetime("2023-9-30 00:00:00")

for (site in unique(wc_dat$Site)){
     dat <- subset(wc_dat_daily_mrd, Site == site)
     p <- get_wc_d_plot(dat, t1, t2)
     print(p)
    }
```

### Relative Difference of Water Content 
From the spatial average for each site/treatment pair (0 Line in black)

```{r plot mean relative difference wc, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
# t1 <- as_datetime("2022-07-01 00:00:00")
# t2 <- as_datetime("2023-4-6 00:00:00")

# for (site in unique(wc_dat$Site)){
#      dat <- subset(wc_dat_daily_mrd, Site == site)
#      dat_mean <-  subset(wc_dat_daily_mrd_mean, Site == site)
#      p <- get_wc_mrd_plot(dat, dat_mean, t1, t2)
#      print(p)
#     }
```


## Temporal mean & standard deviation of relative difference in water content of each sensor 

### Unranked, ordered by Block 
```{r Plot temporal stability by site for DAILY data unRanked, warning = FALSE, echo = FALSE, fig.width=15,fig.height=10}
# for (site in unique(wc_dat_daily_mrd_mean$Site)){
#      dat <- subset(wc_dat_daily_mrd_mean, Site == site)
#      p1 <- get_mrd_bar_plot(dat)
#      p2 <- get_sd_bar_plot(dat)
#      p <- ggarrange(p1, p2, ncol = 2)
#      print(p)
#     }
```

### Ranked
```{r Plot temporal stability by site for DAILY data RANKED, warning = FALSE, echo = FALSE, fig.width=16,fig.height=8}
for (site in unique(wc_dat_daily_mrd_mean$Site)){
     dat <- subset(wc_dat_daily_mrd_mean, Site == site)
     p1 <- get_mrd_bar_plot2(dat)
     p2 <- get_sd_bar_plot2(dat)
     p <- ggarrange(p1, p2, ncol = 2)
     print(p)
    }
```

## Look at the cross correlation strength with depth for each site
```{r Make correplot, echo = FALSE, fig.width=8, fig.height=8}
library(corrplot)


# need to remove one from RR Port 3 that is NA after  mid Aug 22 because it is causing all rows to be removed after that date


# convert wc_dat from tall to wide format
avg_corr <- data.frame(matrix(ncol=3,nrow=0))
colnames(avg_corr) <- c('Name','Port','rowMeans(cm)')
# make a corrplot of each site and depth combination
for (site in unique(wc_dat$Site)){
     for (port in unique(wc_dat$Port)){
          dat <- subset(wc_dat, Site == site & Port == port & Name != 'RR_AD_4_11')
          dat <- dat[!is.na(dat$value),]
          dat_wide <- dcast(dat, datetimeUTC ~ Name, value.var='value', fun.aggregate=mean)
          rownames(dat_wide) <- dat_wide$datetimeUTC
          dat_wide <- na.omit(dat_wide)
          cm <- cor(dat_wide[,-1])

          cm_mean <- as.data.frame(rowMeans(cm))
          cm_mean$Port = port 
          cm_mean$Name <- rownames(cm_mean)
          rownames(cm_mean) <- 1:nrow(cm_mean)
          avg_corr <- rbind(avg_corr,cm_mean)

          corrplot(cm, diag = FALSE, type = 'upper', is.corr = FALSE, col.lim = c(min(min(cm)),max(max(cm))), 
               col = colorRampPalette(c("white", "deepskyblue", "blue4"))(100),
               title = paste0(site,' ',port),  mar=c(0,0,1,0))

          # plot the best correlation 
          p3 <- ggplot() + geom_line(dat_wide, mapping= aes(x=.data[['datetimeUTC']], 
               y=.data[[(names(dat_wide)[which(cm == min(cm), arr.ind =TRUE)[1]+1])]]), color = 'deepskyblue', linewidth = 1, linetype = 'dotted') + 
               geom_text(data = dat_wide %>% filter(datetimeUTC == last(datetimeUTC)), aes(label = names(dat_wide)[which(cm == min(cm), arr.ind =TRUE)[1]+1], 
                                                           x = datetimeUTC+86400*14, 
                                                           y = .data[[(names(dat_wide)[which(cm == min(cm), arr.ind =TRUE)[1]+1])]]), 
                                                           color = 'deepskyblue') + 
               geom_line(dat_wide, mapping = aes(x=.data[['datetimeUTC']], 
               y=.data[[(names(dat_wide)[which(cm == min(cm), arr.ind =TRUE)[2]+1])]]), color = 'deepskyblue', linewidth = 1, linetype = 'dotted') + 
               geom_text(data = dat_wide %>% filter(datetimeUTC == last(datetimeUTC)), aes(label = names(dat_wide)[which(cm == min(cm), arr.ind =TRUE)[2]+1], 
                                                           x = datetimeUTC+86400*14, 
                                                           y = .data[[(names(dat_wide)[which(cm == min(cm), arr.ind =TRUE)[2]+1])]]), 
                                                           color = 'deepskyblue') + 
               geom_line(dat_wide, mapping= aes(x=.data[['datetimeUTC']], 
               y=.data[[(names(dat_wide)[which(cm == max(cm[cm<1]), arr.ind =TRUE)[1]+1])]]), color = 'blue4', linewidth = 1, linetype = 'dotdash') + 
               geom_text(data = dat_wide %>% filter(datetimeUTC == last(datetimeUTC)), aes(label = names(dat_wide)[which(cm == max(cm[cm<1]), arr.ind =TRUE)[1]+1], 
                                                           x = datetimeUTC+86400*14, 
                                                           y = .data[[(names(dat_wide)[which(cm == max(cm[cm<1]), arr.ind =TRUE)[1]+1])]]), 
                                                           color = 'blue4') + 
               geom_line(dat_wide, mapping = aes(x=.data[['datetimeUTC']], 
               y=.data[[(names(dat_wide)[which(cm == max(cm[cm<1]), arr.ind =TRUE)[2]+1])]]), color = 'blue4', linewidth = 1, linetype = 'dotdash') +
               geom_text(data = dat_wide %>% filter(datetimeUTC == last(datetimeUTC)), aes(label = names(dat_wide)[which(cm == max(cm[cm<1]), arr.ind =TRUE)[2]+1], 
                                                           x = datetimeUTC+86400*14, 
                                                           y = .data[[(names(dat_wide)[which(cm == max(cm[cm<1]), arr.ind =TRUE)[2]+1])]]), 
                                                           color = 'blue4') 
          print(p3 + theme_bw() + xlim(c(first(dat_wide$datetimeUTC),last(dat_wide$datetimeUTC)+86400*20)))
          #  ggplot() + geom_line(dat_wide, mapping=aes(x=datetimeUTC,y=`WY_PR-11-12`)) 
          #                geom_line() 

          plot(dat_wide[,which(cm == min(cm), arr.ind =TRUE)[1,1]+1],dat_wide[,which(cm == min(cm), arr.ind =TRUE)[1,2]+1], col = 'deepskyblue', xlim = c(0.05, 0.4), ylim = c(0.05, 0.4))
          points(dat_wide[,which(cm == max(cm[cm<1]), arr.ind =TRUE)[1,1]+1],dat_wide[,which(cm == max(cm[cm <1]), arr.ind =TRUE)[1,2]+1], col = 'blue4')
          lines(c(0.05, 0.4), c(0.05, 0.4), col = 'black')
          title = paste0(site,' ',port)

     }
    }

```

```{r rank the corr,echo=FALSE, warning = FALSE, fig.width=8,fig.height=8}
avg_corr$avg_corr <- avg_corr[,"rowMeans(cm)"]
avg_corr$Site <- sapply(strsplit(avg_corr$Name, "[_|-]"), "[[", 1)
avg_corr$Treatment <- sapply(strsplit(avg_corr$Name, "[_|-]"), "[[", 2)
avg_corr$Block <- sapply(strsplit(avg_corr$Name, "[_|-]"), "[[", 3)
avg_corr$Triangle <- sapply(strsplit(avg_corr$Name, "[_|-]"), "[[", 4)
avg_corr$Bl_Tri <- paste0(avg_corr$Block, "_", avg_corr$Triangle)

for (site in unique(avg_corr$Site)){
     dat <- subset(avg_corr, Site == site)
     p1 <- get_corr_bar_plot2(dat)
     # p2 <- get_sd_bar_plot2(dat)
     # p <- ggarrange(p1, p2, ncol = 2)
     print(p1)
    }
```