---
title: "Water Cycling Graphical Summary"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
```

```{r Load Libraries, echo = FALSE, warning = FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(lubridate)
```

## Introduction
This is a notebook to make some summary plots for the Water Cycling Team to share with Jason Rowntree on 3/20/23

## Datasets
First load the zentra files that were manually downloaded from ZentraCloud (entire record)

```{r Load Zentra Datasets, echo=FALSE}
dir <- "C:\\Users\\kuhlalex\\WorkingDrive\\Data\\" # set the working directory

zentra_loc <- list()
zentra_dat <- list()
zentra_combined <- data.frame(matrix(ncol = 6, nrow = 0))
colnames(zentra_combined ) <- c("Logger", "datetime","Site", "Treatment", "sensor", "water content")

for (file in list.files(paste0(dir, "Zentra\\Manual\\"))){
    file_name <- strsplit(file, "[(|)]")[[1]][2]
    zentra_loc[[file_name]] <- strsplit(file, "[(|)]")[[1]][1]
    print(file_name)
    zentra_dat[[file_name]] <- read.csv(paste0(dir, "Zentra\\Manual\\", file), skip = 2)
    zentra_dat[[file_name]]$Logger <- strsplit(file, "[(|)]")[[1]][2]
    zentra_dat[[file_name]]$Site <- strsplit(file, "[-|_|(|)]")[[1]][1]
    zentra_dat[[file_name]]$Treatment <- strsplit(file, "[-|_|(|)]")[[1]][2]
    zentra_dat[[file_name]]$Block <- strsplit(file, "[-|_|(|)]")[[1]][3]
    zentra_dat[[file_name]]$datetime <- as_datetime(zentra_dat[[file_name]]$Timestamps)

    zentra_temp <- melt(zentra_dat[[file_name]][, c("m3.m3.Water.Content", "m3.m3.Water.Content.1", "m3.m3.Water.Content.2", "Logger", "datetime", "Site", "Treatment", "Block")],
                  na.rm = FALSE,
                  id.vars = c("Logger", "datetime", "Site", "Treatment", "Block"),
                  variable.name = "sensor",
                  value.name = "water content")
    zentra_combined <- rbind(zentra_combined, zentra_temp)
}

head(zentra_combined)
```

Next, load all of the zentra data available through the API (July-November)

```{r function to Load all ZENTRA from API, echo = FALSE, warning = FALSE}
get_zentra <- function(datpath) {
      z <- list()
      for (file in list.files(paste0(dir, datpath))){
            z[[file]] <- read.csv(paste0(dir, datpath, file))
      }

      zentra_api <- do.call(rbind.data.frame, z) # melt the list
      zentra_api$datetimeUTC <- as_datetime(zentra_api$datetime)
      zentra_api$Name <- zentra_api$Site
      zentra_api$Name <- gsub("^AD", "MI-AD", zentra_api$Name) # append MI site ID to these sites
      zentra_api$Name <- gsub("^PR", "MI-PR", zentra_api$Name) # append MI site ID to these sites

      zentra_api$Site <- sapply(strsplit(zentra_api$Name, "[_|-]"), "[[", 1)
      zentra_api$Treatment <- sapply(strsplit(zentra_api$Name, "[_|-]"), "[[", 2)
      zentra_api$Block <- sapply(strsplit(zentra_api$Name, "[_|-]"), "[[", 3)
      zentra_api$Triangle <- sapply(strsplit(zentra_api$Name, "[_|-]"), "[[", 4)

      zentra_api$Port[zentra_api$Port == 5] <- 3 # rename the port 5 sensor as 3
      return(zentra_api)
}
```

```{r Load zentra API data, echo = FALSE}
zentra_WC <- get_zentra("Zentra\\API\\WaterContent\\")
head(zentra_WC)
zentra_WL <- get_zentra("Zentra\\API\\WaterLevel\\")
head(zentra_WL)
```
Load the weather data for Lake City and convert the datetime object

```{r Load Weather Datasets, echo=FALSE, warning = FALSE}
MAWN_dat <- read.csv(paste0(dir, "Weather\\mawn_LakeCity_arl.csv"))
MAWN_dat$datetime <- as_datetime(MAWN_dat$date, format = "%m/%d/%Y")
head(MAWN_dat)
```

Load the combined infiltration data and remove outliers

```{r Load Infiltration Datasets, echo=FALSE}
infil_dat_Fall <- read.csv(paste0(dir, "Infiltration\\Infiltration_combined_Fall.csv"))
infil_dat_Fall$Season <- "Fall"
infil_dat_Spring <- read.csv(paste0(dir, "Infiltration\\Infiltration_combined_Spring.csv"))
infil_dat_Spring$Season <- "Spring"
infil_dat <- rbind(infil_dat_Fall,infil_dat_Spring)
#  clean up NAs
infil_dat[infil_dat == "N/A"] <- NA
infil_dat[, 12] <- as.numeric(infil_dat[, 12])
# summary(infil_dat[, 10:12])
# clean up outliers in KFS data
Q <- quantile(infil_dat[, 10:12], probs = c(0.05, 0.95), na.rm = TRUE)

infil_dat_clean <- infil_dat
infil_dat_clean[which(infil_dat_clean[, 10] < Q[1] | infil_dat_clean[, 10] > Q[2]), 10] <- NA
infil_dat_clean[which(infil_dat_clean[, 11] < Q[1] | infil_dat_clean[, 11] > Q[2]), 11] <- NA
infil_dat_clean[which(infil_dat_clean[, 12] < Q[1] | infil_dat_clean[, 12] > Q[2]), 12] <- NA
# infil_dat_clean$meanKFS <- rowMeans(infil_dat_clean[, 10:12], na.rm = TRUE)
# summary(infil_dat_clean$meanKFS)

head(infil_dat_clean)
infil_dat_tall <- melt(data = infil_dat_clean[, c(2, 4, 6, 7, 10, 11, 12,16)],    
                  na.rm = FALSE,
                  id.vars = c("site", "treatment","triangle","block_name","Season"),
                  variable.name = "experiment",
                  value.name = "KFS_cm/hr")

# infil_dat_gr <- infil_dat %>% group_by(site,treatment) %>% summarise(mean_infil = mean(Kfs.1..cm.hr))
infil_dat_tall$name <- paste0(infil_dat$site, "_", infil_dat$treatment)
infil_dat_tall$triangle <- paste0(infil_dat$site, "_", infil_dat$treatment, "_", infil_dat$block_name, "_", infil_dat$triangle)

```
## Graphics
### Water Content Data 

Plot all of the water content data for the first four months of deployment

* Note that Wyoming sensors were not deployed until September 2022
* In Michigan and Oklahoma, sensors are at 15, 40 and 60 cm depths
* In Wyoming there are at 10, 20, 30 cm due to difficulties augering into the rocky soil
* We observe that the wetter climate of Michigan makes the water content more dynamic
* Water content is quite variable between triangles at the same sites
* No discernable differences between the two treatments at this point

```{r plot water content all, echo = FALSE, warning = FALSE}
t1 <- as_datetime("2022-07-01 0):00:00")
t2 <- as_datetime("2022-11-04 00:00:00")

p1 <- ggplot() + geom_line(data = zentra_WC,
      aes(x = datetimeUTC, y = value, group = interaction(Name,Port), color = as.factor(Port)),
      linewidth = 1, show.legend = FALSE) +
      theme_bw(base_size = 16) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      xlab("Date") + ylab("Water Content (cm^3/cm^3)") +
      xlim(c(t1, t2)) + ylim(c(0, 0.5)) + facet_grid(Treatment ~ factor(Site, levels = c("MI", "CR", "RR", "WY")))
p1
```

* There seems to be an issue with MI-AD_9_1 on Port 1 that is resolved at the time of the Port switch on that logger (z6-18277) on 8/15/22 but that switch moved the sensor on Port 3 to Port 5 - although the issue was actually with Port 1 it was still resolved by the switch. Perhaps it was merely loose?

```{r troubleshoot MI AD, echo = FALSE, warning = FALSE}
t1 <- as_datetime("2022-07-1 01:00:00")
t2 <- as_datetime("2022-11-03 01:00:00")

p2 <- ggplot() + geom_line(data = zentra_WC[zentra_WC$Site == "MI" & zentra_WC$Treatment == "PR",], 
      aes(x = datetimeUTC, y = value, color = as.factor(Port)), 
      linewidth = 1, show.legend = FALSE) +
      theme_bw(base_size = 16) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      xlab("Date") + ylab("Water Content (cm^3/cm^3)") + 
      xlim(c(t1, t2)) + ylim(c(0, 0.5)) + facet_wrap(~Name)
p2
```

* Replot without the noisy data

```{r plot water content all minus noisy sets, echo = FALSE, warning = FALSE}
t1 <- as_datetime("2022-07-12 01:00:00")
t2 <- as_datetime("2022-07-18 01:00:00")
zentra_WC_clean <- zentra_WC[!(zentra_WC$Name == "MI-AD_9_1" & zentra_WC$datetime < as_datetime("2022-8-15 19:00:00")), ]
zentra_WC_clean <- zentra_WC_clean[!(zentra_WC_clean$Name == "MI-PR_9_9" & zentra_WC_clean$Port == 3 &
                   zentra_WC_clean$datetime < as_datetime("2022-8-5 00:00:00") &
                   zentra_WC_clean$datetime > as_datetime("2022-7-28 00:00:00")), ]
zentra_WC_clean <- zentra_WC_clean[!(zentra_WC_clean$Name == "MI-PR_8_11" & zentra_WC_clean$Port == 3 &
                   zentra_WC_clean$datetime < as_datetime("2022-7-16 18:00:00") &
                   zentra_WC_clean$datetime > as_datetime("2022-7-16 12:00:00")), ]


p3 <- ggplot() + geom_line(data = zentra_WC_clean,
      aes(x = datetimeUTC, y = value, group = interaction(Name,Port), color = as.factor(Port)),
      linewidth = 1, show.legend = FALSE) +
      theme_bw(base_size = 16) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      xlab("Date") + ylab("Water Content (cm^3/cm^3)") +
      xlim(c(t1, t2)) + ylim(c(0, 0.5)) + facet_grid(Treatment ~ factor(Site, levels = c("MI", "CR", "RR", "WY")))
p3
```

Do some summary statistics and average each logger through time for each site/treatment pair

```{r plot site/treatment average water content}
zentra_WC_clean$day <- date(zentra_WC_clean$datetimeUTC)
zentra_WC_average <- zentra_WC_clean %>% group_by(Site, Treatment, Port, day) %>% summarise(avg_WC = mean(value))
p4 <- ggplot() + geom_line(data = zentra_WC_average,
      aes(x = day, y = avg_WC, group = as.factor(Port), color = as.factor(Port)),
      linewidth = 1, show.legend = FALSE) +
      theme_bw(base_size = 16) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      xlab("Date") + ylab("Water Content (cm^3/cm^3)") +
      ylim(c(0.05, 0.3)) + facet_grid(Treatment ~ factor(Site, levels = c("MI", "CR", "RR", "WY")))
p4
```


Look at one Logger from a selection of Site/Treatment pairs

* Data shown is from 6 out of 85 measurement locations. There are roughly 3 such locations per plot, and 3 plots per treatment

```{r plot water content at select loggers, echo = FALSE, warning = FALSE}
t1 <- as_datetime("2022-07-01 01:00:00")
t2 <- as_datetime("2022-12-01 01:00:00")

p1 <- ggplot() + geom_line(data = zentra_combined, 
      aes(x = datetime, y = `water content`, color = sensor), size = 1, show.legend = FALSE) +
      theme_bw(base_size = 16) + 
      xlab("Date") + ylab("Water Content (cm^3/cm^3)") + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      xlim(c(t1, t2)) + ylim(c(0,0.4)) + facet_grid(Treatment ~ ~ factor(Site, levels = c("MI", "CR", "RR", "WY")))
p1
```

### Water Content at MI adaptive plot 9 with precip data (units are tenths of cm), July-Nov

* Note increases in soil water content correspond largely with large (> 10 mm) rain events
* Water content increases when ET decreases after the growing season (~October)

```{r plot water content zoom, echo = FALSE, warning = FALSE}
t1 <- as_datetime("2022-07-01 01:00:00")
t2 <- as_datetime("2022-12-01 01:00:00")

p1 <- ggplot() + geom_point(data = zentra_combined[zentra_combined$Logger == "z6-18269",], 
      aes(x = datetime, y = `water content`, col = sensor), show.legend = FALSE) + 
      ggtitle("MI Adaptive, block 9, triangle 7") +
      geom_col(data = MAWN_dat, aes(x = datetime, y = pcpn/100), width = 1, col = "blue") 
p1 + theme_bw(base_size = 16) + xlab("Date") + ylab("Water Content (cm^3/cm^3)") + xlim(c(t1, t2))
```

### Water level data in Michigan

Plot data grouped by treatment

* Note that water levels (indicating recharge) only start to really come up in late fall
* Adaptive treatment seems to have more recharge events happening across all Blocks
* In Block 8 AD evidence of possible topographic gradient targeted by Triangle placement
* Need to account for surface topography & air pressure correction for final interpretation 

```{r plot all water level data, echo = FALSE, warning = FALSE}
t1 <- as_datetime("2022-07-01 01:00:00")
t2 <- as_datetime("2023-03-03 01:00:00")
p2 <- ggplot() + geom_point(data = zentra_WL, 
      aes(x = datetimeUTC, y = value/1000, group = Name, color = Name), 
      alpha = 0.6, size = 1, show.legend = FALSE) +
      theme_bw(base_size = 16) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      xlab("Date") + ylab("Water Level (m)") + 
      xlim(c(t1, t2)) + ylim(c(-0.1, 1.5)) + facet_grid(Treatment ~ Block) #+
      # geom_col(data = MAWN_dat, aes(x = datetime, y = pcpn/10), width = 1, col = "lightblue") 

p2
```


* Zoom in on MI-AD-9-7 
* Note perched water table was dry (Water Level = 0) throughout summer and fall of 2022 and begins to record up to 1.5 m of water above the data logger by Feburary 2023

```{r plot water level at MI-AD-9-7, echo = FALSE, warning = FALSE}
t1 <- as_datetime("2022-7-01 01:00:00")
t2 <- as_datetime("2023-03-01 01:00:00")

p2 <- ggplot() + geom_line(data = zentra_dat[["z6-18269"]], aes(x = datetime, y = mm.Water.Level/1000), size = 2, col = "royalblue") +
      ggtitle("MI Adaptive, block 9, triangle 7") +
     geom_col(data = MAWN_dat, aes(x = datetime, y = pcpn/10), width = 1, col = "lightblue") 
p2 + theme_bw(base_size = 16) + xlab("Date") + ylab("Perched Water Table Level (m)") + xlim(c(t1, t2)) 
```


### Initial infilitration experiment results, grouped by treatment and state

* values outside the 5th and 95th percentile were filtered out
* measurements span February to November 2022, depending on site
* note the much higher values in Michigan and the large variability in measurements

```{r plot infiltration x Season, echo = FALSE, warning = FALSE}
give.n <- function(x){
   return(c(y = median(x) + 1, label = length(x)))
}

p3 <- ggplot(data = infil_dat_tall[!is.na(infil_dat_tall$`KFS_cm/hr`),], aes(x = name, y = `KFS_cm/hr`, fill = site)) +
      geom_boxplot() +
      stat_summary(fun.data = give.n, geom = "text") +
      xlab("Site") + ylab("Infiltration Rate cm/hr") + ggtitle("All Measurements") +
      theme_bw(base_size = 16) +
      theme(axis.text.x=element_text(angle=45,hjust=1)) + facet_wrap(~Season)
p3 + scale_fill_discrete(name = "Intensive Site")      
```

* Note the large variability even within a single ~1 m wide triangle within the same treatment
```{r plot infiltration within triangle, echo = FALSE, warning = FALSE}
p4 <- ggplot(data = infil_dat_tall[infil_dat_tall$name == "OK-RR_PR" & infil_dat_tall$Season == "Fall",],
      aes(x = triangle, y = `KFS_cm/hr`, fill = as.factor(block_name))) +
      geom_boxplot() +
      #stat_summary(fun.data = give.n, geom = "text") +
      xlab("Triangle") + ylab("Infiltration Rate cm/hr") + ggtitle("OK RR Prescriptive") +
      theme_bw(base_size = 16) +
      theme(axis.text.x=element_text(angle=45,hjust=1))
p4 + scale_fill_discrete(name = "Block")
```